---
title: "Epidemic Growth Rates"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(knitr)
```

# Introduction

The R package **incidence** can estimate growth rate and its confidence interval based on: 

* Incidence data, **excluding zeros**, can be modelled using log-linear regression of the form: `log(y) = r x t + b`

* Where y is the incidence, r is the growth rate, t is the number of days since a specific point in time (typically the start of the outbreak), and b is the intercept.

This gives the same result as the coefficient of Base R `glm(log(y) ~ t)`

Additionally you can set the peak or estimate the peak and hence before and after growth rate. 

More info: 

https://f1000research.com/articles/8-139
https://cran.r-project.org/web/packages/incidence/vignettes/overview.html


# Demo

## Incidence package

```{r}
#load packages
library(tidyverse)
library(incidence)
library(readxl)

#import data
x <- read.csv("~/Documents/GitHub/COVID-19/covid_19/growth_rates/scot_cases.csv")
x <- read_excel("/Users/smazeri/Documents/GitHub/COVID-19/covid_19/COVID-19_Scotland_data_all_2020-04-15.xlsx")
path <- "/Users/smazeri/Documents/GitHub/COVID-19/covid_19/COVID-19_Scotland_data_all_2020-04-15.xlsx"
# Scottish cases data
scot_data_raw <- read_excel(path, sheet = "Cases By Health Board", skip = 1)
scot_data_raw <- filter(scot_data_raw, !Health_Board %in% c("Increase", "pIncrease")) %>%
              rename(confirmed_cases= Total,
                     Date = Health_Board) %>%
              mutate(date = lubridate::ymd(Date)) %>%
              select(-Date)
  
scot_data_raw$date[nrow(scot_data_raw)] <- scot_data_raw$date[nrow(scot_data_raw)-1] + days(1)
#scot_data_raw$confirmed_cases[nrow(scot_data_raw)] <- 4565 ## REMOVE

scot_data <- scot_data_raw %>%
             mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases),0)) %>%
             mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) 




#x$date <- lubridate::ymd(x$date)
x <- subset(scot_data, date >= lubridate::ymd(max(scot_data$date)-days(7)))
x <- scot_data
# turn data.frame into incidence object
# can group by location e.g health board
inc <- as.incidence(x = x[,c("new_cases")], dates = x$date)
#set.seed(345)
#inc_b <- bootstrap(inc, randomise_groups = FALSE)
# plot
plot(inc, border = "white") + theme(legend.position = c(0.9, 0.7)) + add_incidence_fit(model_fit_inc)

# fit model
model_fit_inc <- fit(inc)

# growth rate
model_fit_inc$info$r

# growth rate confidence interval
model_fit_inc$info$r.conf
```

```{r}
## Bootstrap results
result <- data.frame(Iteration = NULL,
           `Growth Rate` = NULL,
           LCI_95 = NULL,
           UCI_95 = NULL) 
for (i in 1:1000){
inc_b <- bootstrap(inc, randomise_groups = FALSE)
model_fit_inc <- fit(inc_b)
dat <- data.frame(Iteration = i,
           `Growth Rate` = model_fit_inc$info$r
           #LCI_95 = model_fit_inc$info$r.conf[1],
           #UCI_95 = model_fit_inc$info$r.conf[2]
           ) 

result <- rbind(result, dat)
}

result %>%
  pivot_longer(Growth.Rate:UCI_95, names_to = "Estimate", values_to = "Value") %>%
  ggplot() +
  geom_histogram(aes(x = Value, fill = Estimate)) + facet_grid(Estimate~.)

quantile(result$Growth.Rate, probs = c(0.025, 0.975))

ggplot(x) + geom_bar(aes(x = date, y = log(new_cases)), stat = "identity")
```


## Fit model with base R

```{r}
x$days_since <- x$date - min(x$date)
model_fit <- lm(log(new_cases) ~ days_since, data = subset(x, new_cases >0))
model_fit <- lm(log(confirmed_cases) ~ days_since, data = x)

# growth rate is the coefficient of the variable days_since
model_fit$coefficients[2]
confint(model_fit)[2,]

summary(model_fit)$r.squared
```

## Comparison of results as a sanity check

```{r, echo=FALSE}
data.frame(Method = c("Incidence pachage", "Base R"),
           `Growth Rate` = c(model_fit_inc$info$r, model_fit$coefficients[2][[1]]),
           LCI_95 = c(model_fit_inc$info$r.conf[1], confint(model_fit)[2,][[1]]),
           UCI_95 = c(model_fit_inc$info$r.conf[2], confint(model_fit)[2,][[2]])) %>%
  kable()
```




