---
title: "COVID-19 Scotland daily update"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
output: html_document
---

```{r setup, include=FALSE}
#date: "`r format(Sys.time(), '%d %B, %Y')`" 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Load packages
library(flexdashboard) ; library(shiny) ; library(readr); library(dplyr); library(tidyr); library(purrr); library(forcats); library(stringr); library(htmlwidgets); library(lubridate); library(sf); library(RcppRoll); library(plotly); library(shinythemes);library(leaflet); library(classInt); library(ggrepel); library(scales); library(leaflet.extras); library(RColorBrewer);
library(colorblindr); library(readxl)


# Import Scottish covid data
path <- "COVID-19_Scotland_data_all_2020-04-10.xlsx" 
# Scottish cases data
scot_data_raw <- read_excel(path, sheet = "Cases By Health Board", skip = 1)
scot_data_raw <- filter(scot_data_raw, !Health_Board %in% c("Increase", "pIncrease")) %>%
              rename(confirmed_cases= Total,
                     Date = Health_Board) %>%
              mutate(date = lubridate::ymd(Date)) %>%
              select(-Date)
  
scot_data_raw$date[nrow(scot_data_raw)] <- scot_data_raw$date[nrow(scot_data_raw)-1] + days(1)
#scot_data_raw$confirmed_cases[nrow(scot_data_raw)] <- 4565 ## REMOVE

scot_data <- scot_data_raw %>%
             mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases),0)) %>%
             mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) 
             
# Per health board
scot_data_health_board <- scot_data %>% 
  select(date, Ayrshire:`Dumfries and Galloway`) %>%
  pivot_longer(Ayrshire:`Dumfries and Galloway`,
               names_to = "health_board",
               values_to = "confirmed_cases") %>% 
  group_by(health_board) %>%
  mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases), 0)) %>%
  ungroup() %>%
  replace_na(list(new_cases = 0))

# Per health board for map
# Cumulative incidence
scot_ci_hb <- read_excel(path, sheet = "Cumulative Incidence Grouped") %>%
  slice(nrow(.)) %>%
  select(Ayrshire:Tayside) %>%
  pivot_longer(Ayrshire:Tayside, names_to = "health_board", values_to = "cumulative_incidence")

# Incidence over last day
scot_ti_hb <- read_excel(path, sheet = "Incidence by Health Board") %>%
  slice(nrow(.)) %>%
  select(Ayrshire:Tayside) %>%
  pivot_longer(Ayrshire:Tayside, names_to = "health_board", values_to = "today_incidence")

scot_data_health_board_total <- scot_data_health_board %>% 
                                group_by(health_board) %>%
                                summarise(confirmed_cases = max(confirmed_cases, na.rm = T)) %>%
                                left_join(scot_ci_hb)

# Scottish death data
scot_deaths <- read_excel(path, sheet = "Scotland Deaths", skip = 1) %>%
  rename("deaths" = Deaths_Cum, 
         "new_deaths" = Deaths_New) %>%
  mutate(doubling_time_week = 7*log(2)/log(deaths/replace_na(lag(deaths,7),0))) %>%
  mutate(date = lubridate::ymd(Date)) 

scot_deaths$date[nrow(scot_deaths)] <- scot_deaths$date[nrow(scot_deaths)-1] + days(1)

# Scottish tests
scot_tests <- read_excel(path, sheet = "CPT & DPC") 
#scot_tests$Cases[nrow(scot_tests)] <- 4565 #REMOVE

scot_tests  <- scot_tests %>%
  rename("Conducted" = Tests, 
         "Total Positive" = Cases,
         "deaths_per_case" = DPC,
         "cases_per_test" = CPT) %>%
  mutate("Conducted today" = Conducted - replace_na(lag(Conducted), 0)) %>%
  mutate("Total Negative" = Conducted - `Total Positive`) %>%
  mutate("Positive" = `Total Positive` - replace_na(lag(`Total Positive`), 0),
         "Negative" = `Total Negative` - replace_na(lag(`Total Negative`), 0)) 
scot_tests$Positive[scot_tests$Date == ymd("2020-03-12")] <- 24
scot_tests$Negative[scot_tests$Date == ymd("2020-03-12")] <- 552

  
scot_tests_long <- scot_tests %>%
pivot_longer(cols = Positive:Negative, names_to = "Result", values_to = "Number") %>%
mutate(Result = factor(Result, levels = c("Positive", "Negative")))


# Map files
# SCOTLAND MAP
cases_by_area <- sf::st_read("SG_NHS_HealthBoards_2019b.geojson") %>%
                 mutate(health_board = case_when(HBName == "Ayrshire and Arran" ~ "Ayrshire",
                                                 HBName %in% c("Grampian", "Shetland", "Orkney") ~ "Grampian, Shetland and Orkney",
                                                 HBName %in% c("Highland", "Western Isles") ~ "Highland and Western Isles",
                        TRUE ~ as.character(HBName))) %>%
                 st_transform(crs = st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")) %>%
                 left_join(scot_data_health_board_total, by = c("health_board" = "health_board")) 

# Plot colours
nice_blue <- palette_OkabeIto[5]
nice_red <- palette_OkabeIto[6]  
```

<br>

## Overall update

Here is the daily report for COVID-19 in Scotland for `r weekdays(max(scot_data$date))`, `r format(max(scot_data$date), '%d %B %Y')`. For the full dashboard, including comparisons with other countries, please visit [here](https://smazeri.shinyapps.io/Covid19_Scotland "COVID-19 Scotland").


**Deaths:** <br/>
The death toll is currently `r max(scot_deaths$deaths, na.rm = TRUE)`. <br/>
The number of new deaths reported today was `r scot_deaths$new_deaths[nrow(scot_deaths)]`, which represents a `r formatC(round(100*(scot_deaths$deaths[nrow(scot_deaths)]-scot_deaths$deaths[nrow(scot_deaths)-1])/scot_deaths$deaths[nrow(scot_deaths)-1],1), format='f', digits=1 )`% increase on yesterday's total. <br/>
The doubling time for deaths over the past 7 days was `r formatC(round(scot_deaths$doubling_time_week[nrow(scot_deaths)],1), format='f', digits=1 )` days.

**Cases:** <br/>
The total number of cases is currently `r comma(scot_data$confirmed_cases[nrow(scot_data)])`. <br/> 
The number of new cases today is `r scot_data$new_cases[nrow(scot_data)]`, which represents a `r formatC(round(100*(scot_data$confirmed_cases[nrow(scot_data)]-scot_data$confirmed_cases[nrow(scot_data)-1])/scot_data$confirmed_cases[nrow(scot_data)-1],1), format='f', digits=1 )`% increase on yesterday's total. <br/>
The doubling time for the number of cases over the past 7 days was `r formatC(round(scot_data$doubling_time_week[nrow(scot_data)],1), format='f', digits=1 )` days.

```{r}
data_incidence <- scot_data_health_board %>%
  group_by(health_board) %>%
  arrange(date) %>%
  mutate(relative_increase_cases = 100*(confirmed_cases - replace_na(lag(confirmed_cases),0))/replace_na(lag(confirmed_cases),0)) %>%
  mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) %>%
  ungroup()


data_incidence_today <- subset(data_incidence, date == max(scot_data$date))
```


**Cases per capita (measured per 10,000 head of population):** <br/>
The highest incidence over the last day was `r formatC(round(max(scot_ti_hb$today_incidence),1), format='f', digits=1 )`, which occurred in `r subset(scot_ti_hb, today_incidence == max(scot_ti_hb$today_incidence))$health_board`. <br/>
The highest cumulative incidence is `r formatC(round(max(scot_ci_hb$cumulative_incidence),1), format='f', digits=1 )`, which is in `r subset(scot_ci_hb, cumulative_incidence == max(scot_ci_hb$cumulative_incidence))$health_board`.

**Health Boards\*:** <br/>
The largest increase in cases in absolute terms by health board was `r max(data_incidence_today$new_cases)` cases, which occurred in `r subset(data_incidence_today, new_cases == max(data_incidence_today$new_cases))$health_board`. <br/>
The largest relative increase in cases was `r formatC(round(max(data_incidence_today$relative_increase_cases),1),format='f', digits=1)`%, which occurred in `r subset(data_incidence_today, relative_increase_cases == max(data_incidence_today$relative_increase_cases))$health_board`. <br/>
The fastest doubling time over the past 7 days was `r formatC(round(min(subset(data_incidence_today, doubling_time_week>0)$doubling_time_week),1),format='f', digits=1)` days, which occurred in `r subset(data_incidence_today, doubling_time_week == min(subset(data_incidence_today, doubling_time_week>0)$doubling_time_week))$health_board`.
<!-- format(round(min(subset(data_incidence_today, doubling_time_week>0)$doubling_time_week)),nsmall = 1)-->

**Tests:** <br/>
The total number of tests reported is `r comma(scot_tests$Conducted[nrow(scot_tests)])`. <br/>
The total number of positive cases to date is `r comma(scot_tests[nrow(scot_tests),"Total Positive"][[1]])`. <br/>
The percentage of tests which were positive is currently `r formatC(round(100*scot_tests[nrow(scot_tests),"Total Positive"]/scot_tests$Conducted[nrow(scot_tests)],1)[[1]],format='f', digits=1 )`%.

<br> 

\*To enhance graph readability and to account for cases being reallocated across health boards, the following health boards are grouped together: Grampian, Shetland and Orkney; Highland and Western Isles. 

The data in this report are provided by Health Protection Scotland and published daily by the Scottish Government [here](https://www.gov.scot/coronavirus-covid-19/ "gov.scot/coronavirus-covid-19/"). This report has been compiled in collaboration between members of the [Epigroup](https://www.wiki.ed.ac.uk/display/Epigroup/People "Epigroup website") and [EERA](https://www.ed.ac.uk/roslin/eeragroup "EERA website") groups at the University of Edinburgh. 

<br> 


## Plots
### Confirmed cases
```{r}
# annotations
a <- list(
  text = "Daily confirmed cases",
  xref = "paper",
  yref = "paper",
  yanchor = "bottom",
  xanchor = "center",
  align = "center",
  x = 0.5,
  y = 1,
  showarrow = FALSE
)

b <- list(
  text = "Cumulative confirmed cases",
  xref = "paper",
  yref = "paper",
  yanchor = "bottom",
  xanchor = "center",
  align = "center",
  x = 0.5,
  y = 1,
  showarrow = FALSE
)

fig1 <-plot_ly(data = scot_data, x = ~date, y = ~new_cases, text = ~new_cases, color = ~I(nice_blue), type = "bar", hovertemplate = paste('Date: %{x}',
                    '<br>New cases: %{text}<extra></extra>')) %>%
  config(displayModeBar = FALSE)  %>%
  layout(annotations = a,
         barmode = 'stack',
         xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        title = ""), 
        yaxis = list(title = "")) %>%
    partial_bundle()

fig2 <- plot_ly(data = scot_data, x = ~date, y = ~confirmed_cases, color = ~I(nice_blue), type = "scatter", mode = "lines+markers", 
               hovertemplate = paste('Date: %{x}',
                        '<br>Total cases: %{y}<extra></extra>')) %>%
config(displayModeBar = FALSE)  %>%
  layout(annotations = b,
        xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        title = ""), 
        yaxis = list(title = "",
                     range=c(0,max(scot_data$confirmed_cases) + 5),
                     tickformat = "digit"),
        showlegend = FALSE) %>%
    partial_bundle()

subplot(fig1, fig2, nrows = 2, margin = 0.07)
```

### Deaths
```{r}
# annotations
a <- list(
  text = "Daily confirmed deaths",
  xref = "paper",
  yref = "paper",
  yanchor = "bottom",
  xanchor = "center",
  align = "center",
  x = 0.5,
  y = 1,
  showarrow = FALSE
)

b <- list(
  text = "Cumulative confirmed deaths",
  xref = "paper",
  yref = "paper",
  yanchor = "bottom",
  xanchor = "center",
  align = "center",
  x = 0.5,
  y = 1,
  showarrow = FALSE
)
fig1 <- scot_deaths %>%
    plot_ly(
        x = ~date,
        y = ~new_deaths,
        color = I(nice_red),
        type = "bar",
  hovertemplate = paste('Date: %{x}',
                        '<br>New deaths: %{y}<extra></extra> ')) %>%
  config(displayModeBar = FALSE)  %>%
  layout(annotations = a,
         xaxis = list(
        type = 'date',
        tickformat = "%d %B", title = ""), 
        yaxis = list(title = "")) %>%
    partial_bundle()

fig2<-plot_ly(data = scot_deaths, x = ~date, y = ~deaths, color = ~I(nice_red), type = "scatter", mode = "lines+markers", 
               hovertemplate = paste('Date: %{x}',
                        '<br>Total deaths: %{y}<extra></extra>')) %>%
  config(displayModeBar = FALSE)  %>%
  layout(annotations = b,
         xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        title = ""), 
        yaxis = list(title = "",
                     range=c(0,max(scot_deaths$deaths) + 5),
                     tickformat = "digit"),
        showlegend = FALSE) %>%
    partial_bundle()

subplot(fig1, fig2, nrows = 2, margin = 0.07)
```


### Confirmed cases by Health Board in Scotland
```{r}
ft_fill <- "white"

cov_offset <- function(df, ref = 10, var){
  day_zero <- df %>% 
    arrange(date) %>% 
    filter({cases} >= ref) %>% 
    pull(date) %>% 
    min()
  
  df %>% 
    mutate(days = date - day_zero) %>% 
    filter(days >= 0) %>% 
    mutate(days = as.numeric(days))
}

ref <- 10

ts <- scot_data_raw %>%
  rename("Total" = confirmed_cases) %>%
  pivot_longer(Ayrshire:Total,
               names_to = "health_board",
               values_to = "cases") %>% 
  mutate(health_board = fct_relevel(health_board, "Total", after = Inf)) %>% 
  select(date, health_board, cases) %>% 
  group_by(health_board) %>% 
  nest() %>% 
  mutate(offset = map(data, cov_offset, ref = ref, var = cases)) %>% 
  unnest(offset) %>% 
  select(-data)

line_2 <- tibble( days = seq(0, max(ts$days)-8)) %>% 
  mutate(cases = c(ref, ref * exp(log(2) / 2) ^ (1:(n()-1))))

line_3 <- tibble( days = seq(0, max(ts$days))) %>% 
  mutate(cases = c(ref, ref * 1.260 ^ (1:(n()-1))))

line_7 <- tibble( days = seq(0, max(ts$days))) %>% 
  mutate(cases = c(ref, ref * 1.104 ^ (1:(n()-1))))

p<- ts %>% 
  group_by(health_board) %>% 
  mutate(label = if_else(days == max(days),
                         as.character(health_board), NA_character_)) %>%
  ggplot() +
  aes(x = days, cases, colour = health_board) +
  geom_line(lwd = 0.75, alpha = 0.75) +
  geom_point(size = 1) +
  scale_x_continuous(breaks = function(x) seq(0, round(max(x)), 5)) +
  scale_y_continuous(trans = "log10",
                     breaks = c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000, 5000),
                     labels = scales::comma_format()) +
  #scale_color_manual(values = colorRampPalette(brewer.pal(8,"Spectral"))(12))+
  scale_color_manual(values = c('#e6194b', '#ffe119', '#4363d8','#800000', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#008080', '#9a6324',  '#808000', '#ffd8b1', '#000075', '#808080')) +
  labs(x = str_glue("Days since {ref}th case"),
       y = "Cumulative cases",
       caption = "dotted line: doubling every 3 days; dashed line: doubling every 7 days") +
  #geom_line(data = line_2, aes(x = days, y = cases), colour = "grey30", linetype = "twodash") +
  #geom_line(data = line_3, aes(x = days, y = cases), colour = "grey30", lty = 3) +
  #geom_line(data = line_7, aes(x = days, y = cases), colour = "grey30", lty = 2) +
  theme_bw(base_size = 12) +
  theme(#legend.position = "none",
        plot.background = element_rect(fill = ft_fill),
        panel.background = element_rect(fill = ft_fill),
        panel.grid.major.y = element_line(colour = "grey70",
                                          linetype = "solid",
                                          size = 0.25),
        panel.grid.major.x = element_line(colour = "grey70",
                                          linetype = "solid",
                                          size = 0.25))

w <- ggplotly(p) %>%
  layout(legend = list(
      orientation = "h",
      #position = "top"
      x = 0, 
      y = max(ts$cases)
      )
    ) %>%
    partial_bundle()

#w$x$data

w #%>%
  #style(text = "Doubling in 2 days", traces = 13) %>%
  #style(text = "Doubling in 3 days", traces = 14) %>%
  #style(text = "Doubling in 7 days", traces = 15)
```

### Map

```{r, eval = TRUE}
  breaks <- classIntervals(cases_by_area$confirmed_cases, n = 4, style = "jenks")$brks
  pal <- colorBin(palette = "Blues", domain = NULL, bins = breaks, na.color = "#FFFFFF")
    breaks <- classIntervals(cases_by_area$cumulative_incidence, n = 4, style = "jenks")$brks
  pal2 <- colorBin(palette = "Blues", domain = NULL, bins = breaks, na.color = "#FFFFFF")
  
cases_by_area$cases_popup <- paste0("<b>",cases_by_area$health_board,"</b>","<br/>", cases_by_area$confirmed_cases, " cases")
cases_by_area$cases_popup_pop <- paste0("<b>",cases_by_area$health_board,"</b>","<br/>", round(cases_by_area$cumulative_incidence,2), " cases per", "<br/>", "10,000 population")

labs1 <- as.list(cases_by_area$cases_popup)
labs2 <- as.list(cases_by_area$cases_popup_pop)
  
  leaflet(data = cases_by_area) %>%
    #setView(-3, 54.3, zoom = 5) %>% 
    addTiles(urlTemplate = "", attribution = 'Copyright Scottish Government, contains Ordnance Survey data © Crown copyright and database right (2019) | Data: Health Protection Scotland and Scottish Government', options = providerTileOptions(minZoom = 5, maxZoom = 9)) %>%
    addPolygons(fillColor = ~pal(confirmed_cases), fillOpacity = 0.8, smoothFactor = 0.5, stroke = TRUE, weight = 1, color = "#bdbdbd", opacity = 1, label = lapply(labs1, HTML), labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "15px", direction = "auto"), highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE), group = "Numbers") %>% 
    addPolygons(fillColor = ~pal2(cumulative_incidence), fillOpacity = 0.8, smoothFactor = 0.5, stroke = TRUE, weight = 1, color = "#bdbdbd", opacity = 1, label = lapply(labs2, HTML), labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "15px", direction = "auto"), highlightOptions = highlightOptions(color = "white", weight = 2), group ="Rate") %>%
    addLegend(pal = pal, values = ~confirmed_cases, opacity = 0.7, title = "Cases", position = "bottomright", group = "Numbers") %>% 
    addLegend(pal = pal2, values = ~cumulative_incidence, opacity = 0.7, title = "Rate", position = "bottomright", group = "Rate") %>% 
      addLayersControl(
    #baseGroups = c("Numbers", "Rate"),
    overlayGroups = c("Numbers", "Rate"),
    options = layersControlOptions(collapsed = FALSE)) %>%
    hideGroup("Rate") %>%
    addFullscreenControl() %>% 
    addResetMapButton() %>% 
    onRender(
      "function(el, t) {var myMap = this;myMap._container.style['background'] = '#ffffff';}",
      paste0("function(el, x) {$('head').append(","\'\'",");}"))
```

### Tests conducted
```{r}
plot_ly(data = scot_tests_long, x = ~Date, y = ~Number, text = ~`Conducted today`, color = ~Result, type = "bar", colors = c(nice_blue, "#009E73"), 
               hovertemplate = paste('Date: %{x}',
                        '<br>Result P/N: %{y}',
                        '<br>Total test: %{text}<extra></extra>')) %>%
  config(displayModeBar = FALSE)  %>%
  layout(barmode = 'stack',
         xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        #showgrid = F,
        title = ""), 
        yaxis = list(title = "",
                     tickformat = "digit"),
        legend = list(x = 0.1, y = 0.9)) %>%
    partial_bundle()
```

### Confirmed cases per test conducted
```{r}
plot_ly(data = scot_tests, x = ~Date, y = ~cases_per_test*100, color = ~I(nice_blue), type = "scatter", mode = "lines+markers", 
               hovertemplate = paste('Date: %{x}',
                        '<br>Cases per test: %{y:.2f}<extra></extra>')) %>%
  config(displayModeBar = FALSE)  %>%
  layout(xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        title = ""), 
        yaxis = list(title = "%"),
        showlegend = FALSE) %>%
    partial_bundle()
```

### Deaths per confirmed case
```{r}
plot_ly(data = scot_tests, x = ~Date, y = ~deaths_per_case*100, color = ~I(nice_red), type = "scatter", mode = "lines+markers", 
               hovertemplate = paste('Date: %{x}',
                        '<br>Deaths per case: %{y:.2f}<extra></extra>')) %>%
  config(displayModeBar = FALSE)  %>%
  layout(xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        title = ""), 
        yaxis = list(title = "%"),
        showlegend = FALSE) %>%
    partial_bundle()
```


```{r, eval=FALSE}
### Doubling times
doubling_times_cases_deaths <- bind_rows(select(scot_data, date, doubling_time_week) %>% mutate(Type = "confirmed_cases"), select(scot_deaths, date, doubling_time_week) %>% mutate(Type = "deaths"))

plot_ly(data = doubling_times_cases_deaths, x = ~date, y = ~doubling_time_week, text = ~Type, color = ~Type, type = "scatter", mode = "lines+markers", colors = c(nice_blue, nice_red), 
               hovertemplate = paste('Date: %{x}',
                        '<br>Doubling time: %{y:.2f} days',
                        '<br>Type: %{text}<extra></extra>')) %>%
  #config(displayModeBar = FALSE)  %>%
  layout(xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        #showgrid = F,
        title = ""), 
        yaxis = list(title = "Doubling time over the past 7 days (in days)",
                     autorange='reversed'),
        legend = list(x = 0.1, y = 0.1)) %>%
    partial_bundle()
```


```{r, eval=FALSE}
### Doubling times per health board
plot_ly(data = data_incidence, x = ~date, y = ~doubling_time_week, color = ~health_board, 
    colors = colorRampPalette(brewer.pal(8,"Spectral"))(12),
          text = ~health_board,type = "scatter", mode = "lines+markers", 
               hovertemplate = paste('Date: %{x}',
                        '<br>Doubling time: %{y:.2f}',
                        '<br>Health board: %{text}<extra></extra>')) %>%
  #config(displayModeBar = FALSE)  %>%
  layout(xaxis = list(
        type = 'date',
        tickformat = "%d %B", 
        #showgrid = F,
        title = ""), 
        yaxis = list(title = "Doubling time over the past 7 days (in days)",
                     autorange='reversed'),
        legend = list(orientation = "h",   # show entries horizontally
                     xanchor = "center",  # use center of legend as anchor
                     x = 0.5)) %>%
    partial_bundle()
```


