---
title: "COVID-19 Scotland daily update"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
header-includes:
   - \usepackage{booktabs}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage{caption}
   - \captionsetup[figure]{labelformat=empty}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
urlcolor: blue
---

```{r setup, include=FALSE}
#date: "`r format(Sys.time(), '%d %B, %Y')`" 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Load packages
library(flexdashboard) ; library(shiny) ; library(readr); library(dplyr); library(tidyr); library(purrr); library(forcats); library(stringr); library(htmlwidgets); library(lubridate); library(sf); library(RcppRoll); library(plotly); library(shinythemes);library(leaflet); library(classInt); library(ggrepel); library(scales); library(leaflet.extras); library(RColorBrewer);
library(colorblindr); library(readxl);library(spatstat.utils);library(httr);library(cowplot)

source("dailyreport_dataprep_wk.R")
```



\subsection*{Overall update}

Here is the daily report for COVID-19 in Scotland for `r weekdays(max(scot_data$date))`, `r format(max(scot_data$date), '%d %B %Y')`. For the full dashboard, including comparisons with other countries, please visit [here](https://smazeri.shinyapps.io/Covid19_Scotland "COVID-19 Scotland").


**Deaths:**   
The death toll is currently `r comma(max(scot_deaths$deaths, na.rm = TRUE))`.   
The number of new deaths reported today was `r scot_deaths$new_deaths[nrow(scot_deaths)]`, which represents a `r formatC(round(100*(scot_deaths$deaths[nrow(scot_deaths)]-scot_deaths$deaths[nrow(scot_deaths)-1])/scot_deaths$deaths[nrow(scot_deaths)-1],1), format='f', digits=1 )`% increase on yesterday's total.\newline
The weekly ratio for new deaths measured over the last 7 days was `r formatC(round(wk_gr_newdeaths$ratio_m[nrow(wk_gr_newdeaths)],1), format='f', digits=1 )`. This is `r wk_gr_newdeaths$significance[nrow(wk_gr_newdeaths)]`.  

**Cases:**   
The total number of cases is currently `r comma(scot_data$confirmed_cases[nrow(scot_data)])`.    
The number of new cases today is `r scot_data$new_cases[nrow(scot_data)]`, which represents a `r formatC(round(100*(scot_data$confirmed_cases[nrow(scot_data)]-scot_data$confirmed_cases[nrow(scot_data)-1])/scot_data$confirmed_cases[nrow(scot_data)-1],1), format='f', digits=1 )`% increase on yesterday's total. \newline
The weekly ratio for new cases measured over the last 7 days was `r formatC(round(wk_gr_newcases$ratio_m[nrow(wk_gr_newcases)],1), format='f', digits=1 )`. This is `r wk_gr_newcases$significance[nrow(wk_gr_newcases)]`. 

**Hospital and ICU Occupancy**       
The number of patients in hospital with confirmed or suspected COVID-19 as of midnight is `r comma(filter(data_hosp, date == max(data_hosp$date))$Hospital_total)`, of whom `r filter(data_hosp, date == max(data_hosp$date))$ICU_total` are in ICU.      
The day-of-the-week ratio for hospital occupancy `r formatC(round(wk_gr_hosp_hosp_latest$ratio_m,1), format='f', digits=1 )`. This is `r wk_gr_hosp_hosp_latest$significance`.       
The day-of-the-week ratio for ICU occupancy is `r formatC(round(wk_gr_hosp_icu_latest$ratio_m,1), format='f', digits=1 )`. This is `r wk_gr_hosp_icu_latest$significance`.

```{r}
data_incidence <- scot_data_health_board %>%
  group_by(health_board) %>%
  arrange(date) %>%
  mutate(relative_increase_cases = 100*(confirmed_cases - replace_na(lag(confirmed_cases),0))/replace_na(lag(confirmed_cases),0)) %>%
  mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) %>%
  ungroup()


data_incidence_today <- subset(data_incidence, date == max(scot_data_health_board$date))

data_growth_rates_today <- subset(wk_gr_newcases_hb, date == max(wk_gr_newcases_hb$date)) %>%
                           arrange(desc(ratio_m)) %>% 
                           .[1,]
```


**Cases per capita (measured per 10,000 head of population):**   
The highest incidence over the last day was `r formatC(round(max(scot_ti_hb$today_incidence),1), format='f', digits=1 )`, which occurred in `r subset(scot_ti_hb, today_incidence == max(scot_ti_hb$today_incidence))$health_board`.  
The highest cumulative incidence is `r formatC(round(max(scot_ci_hb$cumulative_incidence),1), format='f', digits=1 )`, which is in `r subset(scot_ci_hb, cumulative_incidence == max(scot_ci_hb$cumulative_incidence))$health_board`.

**Health Boards\*:**   
The largest increase in cases in absolute terms by health board was `r max(data_incidence_today$new_cases)` cases, which occurred in `r subset(data_incidence_today, new_cases == max(data_incidence_today$new_cases))$health_board`.   
The largest relative increase in cases was `r formatC(round(max(data_incidence_today$relative_increase_cases),1),format='f', digits=1)`%, which occurred in `r subset(data_incidence_today, relative_increase_cases == max(data_incidence_today$relative_increase_cases))$health_board`. \newline
The highest weekly ratio for new cases was `r formatC(round(data_growth_rates_today$ratio_m,1), format='f', digits=1)`, which occurred in `r data_growth_rates_today$health_board`. `r data_growth_rates_today$sign_ci`.

**Tests:**   
The total number of tests reported is `r comma(scot_tests$Conducted[nrow(scot_tests)])`.   
The total number of positive cases to date is `r comma(scot_tests[nrow(scot_tests),"Total Positive"][[1]])`.   
The percentage of tests which were positive is currently `r formatC(round(100*scot_tests[nrow(scot_tests),"Total Positive"]/scot_tests$Conducted[nrow(scot_tests)],1)[[1]],format='f', digits=1 )`%.





\*To enhance graph readability and to account for cases being reallocated across health boards, the following health boards are grouped together: Grampian, Shetland and Orkney; Highland and Western Isles. 

The data in this report are provided by Health Protection Scotland and published daily by the Scottish Government [here](https://www.gov.scot/coronavirus-covid-19/ "gov.scot/coronavirus-covid-19/"). This report has been compiled in collaboration between members of the [Epigroup](https://www.wiki.ed.ac.uk/display/Epigroup/People "Epigroup website") and [EERA](https://www.ed.ac.uk/roslin/eeragroup "EERA website") groups at the University of Edinburgh. 

<br> 

\newpage

\subsection*{Plots}
\subsubsection*{Confirmed cases}

```{r, fig.width=5.5, fig.height= 7}
fig1 <- ggplot(scot_data, aes(x = date, y = new_cases)) + geom_bar(stat = "identity", fill = nice_blue, colour = "white") + 
        labs(x="", y = "", subtitle = "Daily confirmed cases") +
        geom_hline(yintercept = 0) + theme_minimal(base_size = 12) +
        scale_x_date(date_breaks = "14 days", labels = date_format("%b %d")) +
        theme(plot.subtitle = element_text(size = 8)) +
        theme(plot.margin = unit(c(0.1,0.4,0.1,0.1), "cm"))

fig2 <- ggplot(scot_data, aes(x = date, y = confirmed_cases)) + geom_line(color = nice_blue) + geom_point(color = nice_blue) +
        labs(x="", y = "", subtitle = "Cumulative confirmed cases") +
        geom_hline(yintercept = 0) + theme_minimal(base_size = 12) +
        scale_x_date(date_breaks = "14 days", labels = date_format("%b %d")) +
        theme(plot.subtitle = element_text(size = 8)) +
        theme(plot.margin = unit(c(0.1,0.4,0.1,0.1), "cm"))

gridExtra::grid.arrange(fig1, fig2, nrow = 2)
```

\subsubsection*{Deaths}

```{r, fig.width=5.5, fig.height=7}
fig1 <- ggplot(scot_deaths, aes(x = date, y = new_deaths)) + geom_bar(stat = "identity", fill = nice_red, colour = "white") + 
        labs(x="", y = "", subtitle = "Daily confirmed deaths") +
        geom_hline(yintercept = 0) + theme_minimal(base_size = 12) +
    scale_x_date(date_breaks = "14 days", labels = date_format("%b %d")) +
        theme(plot.subtitle = element_text(size = 8))  +
        theme(plot.margin = unit(c(0.1,0.4,0.1,0.1), "cm"))

fig2 <- ggplot(scot_deaths, aes(x = date, y = deaths)) + geom_line(color = nice_red) + geom_point(color = nice_red) +
        labs(x="", y = "", subtitle = "Cumulative confirmed deaths") +
        geom_hline(yintercept = 0) + theme_minimal(base_size = 12) +
        scale_x_date(date_breaks = "14 days", labels = date_format("%b %d")) +
        theme(plot.subtitle = element_text(size = 8))  +
        theme(plot.margin = unit(c(0.1,0.4,0.1,0.1), "cm"))

gridExtra::grid.arrange(fig1, fig2, nrow = 2)
```


\subsubsection*{Cumulative confirmed cases by Health Board in Scotland}

```{r, fig.width=6, eval = FALSE}
png(filename = "CumCasesHB_16_04_2020.png", width = 6, height = 4,units = "in", res = 400)
ft_fill <- "white"

cov_offset <- function(df, ref = 10, var){
  day_zero <- df %>% 
    arrange(date) %>% 
    filter({cases} >= ref) %>% 
    pull(date) %>% 
    min()
  
  df %>% 
    mutate(days = date - day_zero) %>% 
    filter(days >= 0) %>% 
    mutate(days = as.numeric(days))
}

ref <- 10

ts <- scot_data_raw %>%
  rename("Total" = confirmed_cases) %>%
  pivot_longer(Ayrshire:Total,
               names_to = "health_board",
               values_to = "cases") %>% 
  mutate(health_board = fct_relevel(health_board, "Total", after = Inf)) %>% 
  select(date, health_board, cases) %>% 
  group_by(health_board) %>% 
  nest() %>% 
  mutate(offset = map(data, cov_offset, ref = ref, var = cases)) %>% 
  unnest(offset) %>% 
  select(-data) %>%
  filter(health_board != "Total")

p<- ts %>% 
  group_by(health_board) %>% 
  mutate(label = if_else(days == max(days),
                         as.character(health_board), NA_character_)) %>%
  ggplot() +
  aes(x = days, cases, colour = health_board) +
  geom_line(lwd = 0.75, alpha = 0.75) +
  geom_point(size = 1) +
  scale_x_continuous(breaks = function(x) seq(0, round(max(x)), 5)) +
  scale_y_continuous(trans = "log10",
                     breaks = c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000, 5000),
                     labels = scales::comma_format()) +
  guides(color = guide_legend(nrow = 4)) +
  scale_color_manual(values = c('#e6194b', '#ffe119', '#4363d8','#800000', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#008080', '#9a6324',  '#808000', '#ffd8b1', '#000075'#, '#808080'
                                )) +
  labs(x = str_glue("Days since {ref}th case"),
       y = "Cumulative Cases"
       ) +
  theme_bw(base_size = 10) +
  theme(legend.position="top",
        legend.box="horizontal", legend.margin=margin(),
        legend.title = element_blank(),
        plot.background = element_rect(fill = ft_fill),
        panel.background = element_rect(fill = ft_fill),
        panel.grid.major.y = element_line(colour = "grey70",
                                          linetype = "solid",
                                          size = 0.25),
        panel.grid.major.x = element_line(colour = "grey70",
                                          linetype = "solid",
                                          size = 0.25))

p
dev.off()
```


```{r, fig.width=6, eval = TRUE}
ft_fill <- "white"

cov_offset <- function(df, ref = 10, var){
  day_zero <- df %>% 
    arrange(date) %>% 
    filter({cases} >= ref) %>% 
    pull(date) %>% 
    min()
  
  df %>% 
    mutate(days = date - day_zero) %>% 
    filter(days >= 0) %>% 
    mutate(days = as.numeric(days))
}

ref <- 10

ts <- scot_data_raw %>%
  rename("Total" = confirmed_cases) %>%
  pivot_longer(Ayrshire:Total,
               names_to = "health_board",
               values_to = "cases") %>% 
  mutate(health_board = fct_relevel(health_board, "Total", after = Inf)) %>% 
  select(date, health_board, cases) %>% 
  group_by(health_board) %>% 
  nest() %>% 
  mutate(offset = map(data, cov_offset, ref = ref, var = cases)) %>% 
  unnest(offset) %>% 
  select(-data)

p<- ts %>% 
  group_by(health_board) %>% 
  mutate(label = if_else(days == max(days),
                         as.character(health_board), NA_character_)) %>%
  ggplot() +
  aes(x = days, cases, colour = health_board) +
  geom_line(lwd = 0.75, alpha = 0.75) +
  geom_point(size = 1) +
  scale_x_continuous(breaks = function(x) seq(0, round(max(x)), 5)) +
  scale_y_continuous(trans = "log10",
                     breaks = c(1, 2, 5, 10, 20, 50, 100, 200, 500, 1000, 5000),
                     labels = scales::comma_format()) +
  guides(color = guide_legend(nrow = 4)) +
  scale_color_manual(values = c('#e6194b', '#ffe119', '#4363d8','#800000', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#008080', '#9a6324',  '#808000', '#ffd8b1', '#000075', '#808080')) +
  labs(x = str_glue("Days since {ref}th case"),
       y = "Cumulative cases"
       ) +
  theme_bw(base_size = 12) +
  theme(legend.position="top",
        legend.box="horizontal", legend.margin=margin(),
        legend.title = element_blank(),
        plot.background = element_rect(fill = ft_fill),
        panel.background = element_rect(fill = ft_fill),
        panel.grid.major.y = element_line(colour = "grey70",
                                          linetype = "solid",
                                          size = 0.25),
        panel.grid.major.x = element_line(colour = "grey70",
                                          linetype = "solid",
                                          size = 0.25))

p
```

\newpage 


```{r, fig.width=9, fig.align="center"}
minimal_theme <- theme_minimal() +
     theme(axis.line=element_blank(),
      axis.text=element_blank(),
      axis.ticks=element_blank(),
      axis.title=element_blank(),
      panel.border = element_blank(),
     panel.grid = element_blank(),
     panel.background = element_blank())

map1 <- ggplot(data = cases_by_area) +
    geom_sf(aes(fill = confirmed_cases)) +
    scale_fill_viridis_c(option = "inferno", trans = "sqrt", direction = -1) +
  #scale_fill_gradient(low = "yellow", high = "red", na.value = NA, trans = "sqrt") +
  minimal_theme + labs(fill = "Total cases") + labs(title = "Total COVID-19 Cases in Scotland")

map1
```

\hfill \break

```{r, fig.width=9, fig.align="center"}
map2 <- ggplot(data = cases_by_area) +
    geom_sf(aes(fill = cumulative_incidence)) +
    scale_fill_viridis_c(option = "inferno", direction = -1) +
  minimal_theme + labs(fill = "Cases per \n10,000 people") + labs(title = "Per Capita COVID-19 Cases in Scotland")

map2

#gridExtra::grid.arrange(map1, map2, nrow = 2, )

#Four options are available: "magma" (or "A"), "inferno" (or "B"), "plasma" (or "C"), "viridis" (or "D", the default option) and "cividis" (or "E").
```


```{r, eval = FALSE}
  breaks <- classIntervals(cases_by_area$confirmed_cases, n = 4, style = "jenks")$brks
  pal <- colorBin(palette = "Blues", domain = NULL, bins = breaks, na.color = "#FFFFFF")
    breaks <- classIntervals(cases_by_area$cumulative_incidence, n = 4, style = "jenks")$brks
  pal2 <- colorBin(palette = "Blues", domain = NULL, bins = breaks, na.color = "#FFFFFF")
  
cases_by_area$cases_popup <- paste0("<b>",cases_by_area$health_board,"</b>","<br/>", cases_by_area$confirmed_cases, " cases")
cases_by_area$cases_popup_pop <- paste0("<b>",cases_by_area$health_board,"</b>","<br/>", round(cases_by_area$cumulative_incidence,2), " cases per", "<br/>", "10,000 population")

labs1 <- as.list(cases_by_area$cases_popup)
labs2 <- as.list(cases_by_area$cases_popup_pop)
  
  leaflet(data = cases_by_area) %>%
    #setView(-3, 54.3, zoom = 5) %>% 
    addTiles(urlTemplate = "", attribution = 'Copyright Scottish Government, contains Ordnance Survey data © Crown copyright and database right (2019) | Data: Health Protection Scotland and Scottish Government', options = providerTileOptions(minZoom = 5, maxZoom = 9)) %>%
    addPolygons(fillColor = ~pal(confirmed_cases), fillOpacity = 0.8, smoothFactor = 0.5, stroke = TRUE, weight = 1, color = "#bdbdbd", opacity = 1, label = lapply(labs1, HTML), labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "15px", direction = "auto"), highlightOptions = highlightOptions(color = "white", weight = 2, bringToFront = TRUE), group = "Numbers") %>% 
    addPolygons(fillColor = ~pal2(cumulative_incidence), fillOpacity = 0.8, smoothFactor = 0.5, stroke = TRUE, weight = 1, color = "#bdbdbd", opacity = 1, label = lapply(labs2, HTML), labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "15px", direction = "auto"), highlightOptions = highlightOptions(color = "white", weight = 2), group ="Rate") %>%
    addLegend(pal = pal, values = ~confirmed_cases, opacity = 0.7, title = "Cases", position = "bottomright", group = "Numbers") %>% 
    addLegend(pal = pal2, values = ~cumulative_incidence, opacity = 0.7, title = "Rate", position = "bottomright", group = "Rate") %>% 
      addLayersControl(
    #baseGroups = c("Numbers", "Rate"),
    overlayGroups = c("Numbers", "Rate"),
    options = layersControlOptions(collapsed = FALSE)) %>%
    hideGroup("Rate") %>%
    #addFullscreenControl() %>% 
    #addResetMapButton() %>% 
    onRender(
      "function(el, t) {var myMap = this;myMap._container.style['background'] = '#ffffff';}",
      paste0("function(el, x) {$('head').append(","\'\'",");}"))
```

\subsubsection*{Tests conducted}
```{r}
scot_tests_long$Result <- factor(scot_tests_long$Result, levels = c("Negative", "Positive"))

ggplot(scot_tests_long, aes(x = ymd(Date), y = Number, fill = Result)) + geom_bar(stat = "identity") +
  scale_fill_manual(values = c("#009E73",nice_blue)) + theme_minimal(base_size = 12) +
  scale_x_date(date_breaks = "14 days", labels = date_format("%b %d")) +
  labs(fill = "", y = "", x = "") +
  geom_hline(yintercept = 0) +
  theme(legend.position = c(0.2, 0.8))  +
  theme(plot.margin = unit(c(0.1,0.4,0.1,0.1), "cm"))
```

\newpage

\subsubsection*{Confirmed cases per test conducted}
```{r, fig.width=5, fig.height=3}
ggplot(scot_tests, aes(x = ymd(Date), y = cases_per_test*100)) + geom_line(color = nice_blue) + geom_point(color = nice_blue) +
scale_x_date(date_breaks = "14 days", labels = date_format("%b %d")) +
labs(x="", y = "%") +
geom_hline(yintercept = 0) + theme_minimal(base_size = 12)  +
        theme(plot.margin = unit(c(0.1,0.4,0.1,0.1), "cm"))
```


\subsubsection*{Deaths per confirmed case}

```{r, fig.width=5, fig.height=3}
ggplot(scot_tests, aes(x = ymd(Date), y = deaths_per_case*100)) + geom_line(color = nice_red) + geom_point(color = nice_red) +
scale_x_date(date_breaks = "14 days", labels = date_format("%b %d")) +  
labs(x="", y = "%") +
geom_hline(yintercept = 0) + theme_minimal(base_size = 12)  +
        theme(plot.margin = unit(c(0.1,0.4,0.1,0.1), "cm"))
```

\newpage 

\subsubsection*{Rates of Change}

```{r casesdeathsratios, fig.width=7, fig.height= 7, eval = FALSE}
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End date of current week",
        y = "Weekly ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New cases") +
  theme_bw(base_size = 12) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 11, b = 0, l = 0)))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-6)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 10)

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.52, y = .49, width = .46, height = .4)

fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "End date of current week",
        y = "Weekly ratio",
        subtitle = "New deaths") +
  theme_bw(base_size = 12)

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-6)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10)

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.52, y = .49, width = .46, height = .40)


gridExtra::grid.arrange(fig1, fig2, nrow = 2) 
```

\thispagestyle{empty}

```{r deathscaseshospicu, fig.width=5.5, fig.height= 9, fig.cap="Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 scale. Day-of-the-week ratios for hospital occupancy and ICU occupancy are ratios of daily counts, measured 7 days apart and plotted on a log10 scale. Figure insets highlight more recent ratios and trends."}
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New Cases (Source: HPS)") +
  theme_bw(base_size = 10) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 14, b = 0, l = 0)))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-6)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.57, y = .53, width = .40, height = .3)

fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 10)

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-6)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"),expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.57, y = .53, width = .40, height = .3)


fig3 <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "ICU Occupancy (Source: NHS Boards)") +
  theme_bw(base_size = 10) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 7, b = 0, l = 0)))



fig3_inset <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_icu$date)-6)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig3 <- ggdraw() +
  draw_plot(fig3) +
  draw_plot(fig3_inset, x = 0.57, y = .53, width = .40, height = .3)

fig4 <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Hospital Occupancy (Source: NHS Boards)") +
  theme_bw(base_size = 10) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 14, b = 0, l = 0)))


fig4_inset <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_hosp$date)-6)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig4 <- ggdraw() +
  draw_plot(fig4) +
  draw_plot(fig4_inset, x = 0.57, y = 0.53, width = 0.40, height = 0.3)

gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 4) 
```


\newpage

```{r, fig.width=6}
p<-ggplot(wk_gr_newcases_hb) +
  aes(
    x = date, y = ratio,
    #ymin = lci, ymax = uci,
    colour = health_board, fill = health_board
  ) +
  #geom_errorbar(width = 0, colour = "gray30", alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line() +
  geom_point(aes(shape = symbol), size = 2) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_color_manual(values = c('#e6194b', '#ffe119', '#4363d8','#800000', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#008080', '#9a6324',  '#808000', '#ffd8b1', '#000075', '#808080'))+
  scale_fill_manual(values = c('#e6194b', '#ffe119', '#4363d8','#800000', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#008080', '#9a6324',  '#808000', '#ffd8b1', '#000075', '#808080'))+
  scale_shape_manual(values = c(1,16)) +
  labs( x = "End date of current week",
        y = "Weekly ratio",
        subtitle = "New cases") +
  guides(colour = guide_legend(override.aes = list(shape = NA), nrow = 6),
         shape = guide_legend(nrow = 6)
         ) +
  theme_bw(base_size = 12) +
    theme(legend.position="top",
        legend.box="horizontal", legend.margin=margin(),
        legend.title = element_blank(),
        legend.text=element_text(size=8)) 

p
```

