---
title: "Monitoring the Covid-19 Pandemic in Scotland"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
header-includes:
   - \usepackage{booktabs}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \floatplacement{table}{H}
   - \usepackage{caption}
   - \captionsetup[figure]{labelformat=empty}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
#classoption: landscape
---


```{r setup, include=FALSE}
#date: "`r format(Sys.time(), '%d %B, %Y')`" 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.pos = 'H')
# Load packages
library(flexdashboard) ; library(shiny) ; library(readr); library(dplyr); library(tidyr); library(purrr); library(forcats); library(stringr); library(htmlwidgets); library(lubridate); library(sf); library(RcppRoll); library(plotly); library(shinythemes);library(leaflet); library(classInt); library(ggrepel); library(scales); library(leaflet.extras); library(RColorBrewer);library(toOrdinal);library(knitr);library(kableExtra);library(RcppRoll);
library(colorblindr); library(readxl);library(spatstat.utils);library(httr);library(cowplot); library(viridis);
library(patchwork)

#source("growth_rate_window.R")
source("weekly_ratios_2.R")
weekly_ratios <- weekly_ratios_2
#source("weekly_ratios_byday.R")

# Import healthboard pop data 
hb_pop <- read_csv("Scotland_Population_Size_By_HealthBoard.csv") %>%
  select(Health_Board, Pop_Size) %>%
  slice(1:14) %>%
  mutate(health_board = str_replace(Health_Board, "and ", "& ")) %>%
  mutate(health_board = recode(health_board, "Ayrshire" = "Ayrshire & Arran"))

# Import HB data
url_hps_hb <- "https://www.gov.scot/binaries/content/documents/govscot/publications/statistics/2020/04/coronavirus-covid-19-trends-in-daily-data/documents/covid-19-data-by-nhs-board/covid-19-data-by-nhs-board/govscot%3Adocument/COVID--19%2Bdata%2Bby%2BNHS%2BBoard%2B16%2BMay%2B2020.xlsx?forceDownload=true"
GET(url_hps_hb, write_disk(tmp <- tempfile(fileext = ".xlsx")))
GET(url_hps_hb, write_disk("hps_daily_hb.xlsx", overwrite = TRUE))
hb_cases <- read_excel(tmp, sheet = "Table 1 - Cumulative cases")
hb_icu <- read_excel(tmp, sheet = "Table 2 - ICU patients")
hb_hosp_conf <- read_excel(tmp, sheet = "Table 3a - Hospital Confirmed")
hb_hosp_susp <- read_excel(tmp, sheet = "Table 3b- Hospital Suspected")

# Import hospital data
## Covid daily hospital data
myurl <- "https://www.gov.scot/binaries/content/documents/govscot/publications/statistics/2020/04/coronavirus-covid-19-trends-in-daily-data/documents/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/govscot%3Adocument/Trends%2Bin%2Bdaily%2BCOVID-19%2Bdata%2B-%2B110520.xlsx?forceDownload=true"
#myurl <- "https://www.gov.scot/binaries/content/documents/govscot/publications/statistics/2020/04/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/documents/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/govscot%3Adocument/Data%2BTable%2B%252810-04-2020%2529.xlsx?forceDownload=true"
GET(myurl, write_disk(tmp <- tempfile(fileext = ".xlsx")))
GET(myurl, write_disk("hps_daily.xlsx", overwrite = TRUE))

#data_hosp <- read_excel(tmp, sheet = "Table 1", skip = 3)
data_hosp <- read_excel(tmp, sheet = "Table 2 - Hospital Care", skip = 3)
#data_hosp <- filter(data_hosp, `...1` != Sys.Date()) # CHANGE

data_hosp <- data_hosp %>%
  rename("date" = "...1",
         "ICU_confirmed" = "Confirmed...2",
         "ICU_suspected" ="Suspected...3", 
         "ICU_total" = "Total...4",
         "Hospital_confirmed" = "Confirmed...5",
         "Hospital_suspected" = "Suspected...6",
         "Hospital_total" = "Total...7" ) %>%
  filter(!is.na(ICU_total)) %>%
  mutate(date = lubridate::ymd(date)) %>%
  mutate(ICU_confsusp = case_when(is.na(ICU_confirmed) & is.na(ICU_suspected) ~ ICU_total)) %>%
  mutate(Hospital_confsusp = case_when(is.na(Hospital_confirmed) & is.na(Hospital_suspected) ~ Hospital_total)) %>%
  select(date, contains("ICU"), everything()) %>%
  mutate(date = date - days(1))

data_hosp_icu <- data_hosp %>%
  select(date, contains("ICU")) %>%
  pivot_longer(cols = c(ICU_confirmed, ICU_suspected, ICU_confsusp), names_to = "Patients", values_to = "Number") %>%
  mutate(Patients = recode(Patients,  "ICU_confsusp" = "Confirmed/Suspected", 
                           "ICU_confirmed" = "Confirmed", 
                           "ICU_suspected" = "Suspected")) %>%
  mutate(Patients = factor(Patients, levels = c("Confirmed/Suspected", "Suspected", "Confirmed")))



data_hosp_hosp <- data_hosp %>%
  select(date, contains("hospital")) %>%
  pivot_longer(cols = c(Hospital_confirmed, Hospital_suspected, Hospital_confsusp), names_to = "Patients", values_to = "Number") %>%
  mutate(Patients = recode(Patients,  "Hospital_confsusp" = "Confirmed/Suspected", 
                           "Hospital_confirmed" = "Confirmed", 
                           "Hospital_suspected" = "Suspected")) %>%
  mutate(Patients = factor(Patients, levels = c("Confirmed/Suspected", "Suspected", "Confirmed")))

wk_gr_hosp_icu <- data_hosp %>%
       mutate(ICU_prev = lag(ICU_total , 7), 
         weekdays(date)) %>%
       filter(date >= ymd("2020-03-25")) %>%
       #pivot_longer(c(ICU_total, ICU_prev), names_to = "week", values_to = "number") %>%
      group_by(date) %>%
      nest() %>%
      mutate(tab = map(data, ~ weekly_ratios_ci(.x$ICU_total, .x$ICU_prev))) %>%
  mutate(ratio_m = map_dbl(tab, ~parse_number(as.character((.x[1,2]))))) %>% 
  mutate(lci = map_dbl(tab, ~parse_number(as.character((.x[2,2]))))) %>% 
  mutate(uci = map_dbl(tab, ~parse_number(as.character((.x[3,2]))))) %>% 
  ungroup() %>%
  select(date, ratio_m, lci, uci) %>%
  mutate(comparison = if_else(ratio_m >1, "greater than", "less than")) %>%
  significance_wk() %>%
  mutate(ratio_under_1 = ratio_m<1)


wk_gr_hosp_icu_latest <- filter(wk_gr_hosp_icu, date == max(wk_gr_hosp_icu$date))

wk_gr_hosp_hosp <- data_hosp %>%
       mutate(Hosp_prev = lag(Hospital_confirmed , 7), 
         weekdays(date)) %>%
       filter(date >= ymd("2020-03-25")) %>%
       #pivot_longer(c(Hospital_total, Hosp_prev), names_to = "week", values_to = "number") %>%
      group_by(date) %>%
      nest() %>%
      mutate(tab = map(data, ~ weekly_ratios_ci(.x$Hospital_confirmed, .x$Hosp_prev))) %>%
  mutate(ratio_m = map_dbl(tab, ~parse_number(as.character((.x[1,2]))))) %>% 
  mutate(lci = map_dbl(tab, ~parse_number(as.character((.x[2,2]))))) %>% 
  mutate(uci = map_dbl(tab, ~parse_number(as.character((.x[3,2]))))) %>% 
  ungroup() %>%
  select(date, ratio_m, lci, uci) %>%
  mutate(comparison = if_else(ratio_m >1, "greater than", "less than")) %>%
  significance_wk() %>%
  mutate(ratio_under_1 = ratio_m<1)

# table(filter(wk_gr_hosp_hosp, date >= max(ymd(wk_gr_hosp_hosp$date))-20)$ratio_under_1)
# table(filter(wk_gr_hosp_hosp, date >= max(ymd(wk_gr_hosp_hosp$date))-20)$ratio_under_1)
# 
# table(filter(wk_gr_hosp_hosp, date >= max(ymd(wk_gr_hosp_hosp$date))-20)$significance)
# table(filter(wk_gr_hosp_hosp, date >= max(ymd(wk_gr_hosp_hosp$date))-20)$ratio_under_1)

# table(filter(wk_gr_hosp_icu, date >= max(ymd(wk_gr_hosp_icu$date))-20)$significance)
# table(filter(wk_gr_hosp_icu, date >= max(ymd(wk_gr_hosp_icu$date))-20)$ratio_under_1)

wk_gr_hosp_hosp_latest <- filter(wk_gr_hosp_hosp, date == max(wk_gr_hosp_icu$date))

# Import NRS data
myurl_deaths <- "https://www.nrscotland.gov.uk/files//statistics/covid19/covid-deaths-data-week-22.xlsx"
GET(myurl_deaths, write_disk(tmp <- tempfile(fileext = ".xlsx")))
GET(myurl_deaths, write_disk("nrs.xlsx", overwrite = TRUE))

#NRS weekly
nrs_deaths_covid_raw <- read_excel(tmp, sheet = "Table 1 - COVID deaths", skip = 3) %>%
  select(-`Year to Date`) %>%
  rename("Details" = "...2",
         "Total" = "...26")

nrs_deaths_covid <- nrs_deaths_covid_raw %>% filter(`Week beginning` == "Deaths involving COVID-194") %>%
  pivot_longer(cols = `43829`:names(nrs_deaths_covid_raw)[ncol(nrs_deaths_covid_raw)-1],
               names_to = "week_beginning", 
               values_to = "weekly_deaths_total") %>%
  mutate(week_beginning = janitor::excel_numeric_to_date(parse_number(week_beginning))) %>%
  mutate(Type = "Covid19 deaths") %>%
  select(week_beginning, weekly_deaths_total, Type)

# NRS daily
nrs_deaths_covid_raw_daily <- read_excel(tmp, sheet = "Figure 2 data", skip =2) %>%
  filter(!is.na(Source)) %>%
  mutate(date = janitor::excel_numeric_to_date(parse_number(Date1))) %>%
  filter(Source == "NRS") %>%
  mutate(new_deaths = c(0,diff(`Cumulative Count`, 1))) 

#Holidays to adjust
# Monday/Tuesday death numbers that can be adjusted are: 13/14 April, 4/5 May, 18/19 May (Victoria day), and for the future 25/26 May (Whitsun). We may also see an effect for 8/9 June (Queen’s official birthday)
#The will affect ratios for 13/20/27 April, 4/11/18/25 May and 1/8 June. 

bank_hol <- dmy(c("13/4/2020", "14/4/2020","04/05/2020", "05/05/2020", "18/05/2020", "19/05/2020", "25/05/2020", "26/05/2020"))

# NRS daily adjust bank holiday effect
nrs_deaths_covid_raw_daily <- nrs_deaths_covid_raw_daily %>%
  mutate(dow = weekdays(date)) %>%
  mutate(bank_hol = if_else(date %in% bank_hol, "Yes", "No"))

nrs_deaths_covid_raw_daily_montues <- nrs_deaths_covid_raw_daily %>%
  filter(dow %in% c("Monday", "Tuesday"))

nrs_non_hol_perc <- nrs_deaths_covid_raw_daily_montues %>% 
  filter(bank_hol == "No") %>%
  group_by(dow) %>%
  summarise(new_deaths_tot = sum(new_deaths)) %>%
  mutate(perc = new_deaths_tot/sum(new_deaths_tot))
  
nrs_deaths_covid_raw_daily_adj <- nrs_deaths_covid_raw_daily_montues %>% 
  filter(bank_hol == "Yes") %>%
  mutate(week_no = week(date)) %>%
  group_by(week_no) %>%
  mutate(new_deaths_tot = sum(new_deaths)) %>%
  mutate(new_deaths_adj = case_when(dow == "Monday" ~ new_deaths_tot*nrs_non_hol_perc$perc[1], 
                                   dow == "Tuesday" ~ new_deaths_tot*nrs_non_hol_perc$perc[2])) %>%
  ungroup() %>%
  select(date, new_deaths_adj)

nrs_deaths_covid_raw_daily <- left_join(nrs_deaths_covid_raw_daily, nrs_deaths_covid_raw_daily_adj, by = "date") %>%
  mutate(new_deaths_adj = coalesce(new_deaths_adj,new_deaths)) 



# Import Scottish covid data
#path <- "COVID-19_Scotland_data_all_2020-04-10.xlsx" 
path <- paste0("Daily_Reports/COVID-19_Scotland_data_all_", {Sys.Date()}, ".xlsx") 
# Scottish cases data
scot_data_raw <- read_excel(path, sheet = "Cases By Health Board", skip = 1)
scot_data_raw <- filter(scot_data_raw, !Health_Board %in% c("Increase", "pIncrease")) %>%
              rename(confirmed_cases= Total,
                     Date = Health_Board) %>%
              mutate(date = lubridate::ymd(Date)) %>%
              select(-Date) 
  
scot_data_raw$date[nrow(scot_data_raw)] <- scot_data_raw$date[nrow(scot_data_raw)-1] + days(1)
#scot_data_raw$confirmed_cases[nrow(scot_data_raw)] <- 4565 ## REMOVE

scot_data_raw <- scot_data_raw %>%
              mutate(date = date - days(1))

scot_data <- scot_data_raw %>%
             mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases),0)) #%>%
             #mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) 

# HPS cases adjust bank holiday effect
scot_data <- scot_data %>%
  mutate(dow = weekdays(date)) %>%
  mutate(bank_hol = if_else(date %in% bank_hol, "Yes", "No"))

scot_data_montues <- scot_data %>%
  filter(dow %in% c("Monday", "Tuesday"))

cases_non_hol_perc <- scot_data_montues %>% 
  filter(bank_hol == "No") %>%
  group_by(dow) %>%
  summarise(new_cases_tot = sum(new_cases)) %>%
  mutate(perc = new_cases_tot/sum(new_cases_tot))
  
scot_data_adj <- scot_data_montues %>% 
  filter(bank_hol == "Yes") %>%
  mutate(week_no = week(date)) %>%
  group_by(week_no) %>%
  mutate(new_cases_tot = sum(new_cases)) %>%
  mutate(new_cases_adj = case_when(dow == "Monday" ~ new_cases_tot*cases_non_hol_perc$perc[1], 
                                   dow == "Tuesday" ~ new_cases_tot*cases_non_hol_perc$perc[2])) %>%
  ungroup() %>%
  select(date, new_cases_adj)

scot_data <- left_join(scot_data, scot_data_adj, by = "date") %>%
  mutate(new_cases_adj = coalesce(new_cases_adj,new_cases)) 

             
# Scottish death data
scot_deaths <- read_excel(path, sheet = "Scotland Deaths", skip = 1) %>%
  rename("deaths" = Deaths_Cum, 
         "new_deaths" = Deaths_New) %>%
  #mutate(doubling_time_week = 7*log(2)/log(deaths/replace_na(lag(deaths,7),0))) %>%
  mutate(date = lubridate::ymd(Date))
  

scot_deaths$date[nrow(scot_deaths)] <- scot_deaths$date[nrow(scot_deaths)-1] + days(1)

scot_deaths <- scot_deaths %>%
  mutate(date = date - days(1)) %>%
  select(-Date)

# HPS deaths adjust bank holiday effect
scot_deaths <- scot_deaths %>%
  mutate(dow = weekdays(date)) %>%
  mutate(bank_hol = if_else(date %in% bank_hol, "Yes", "No"))

scot_deaths_montues <- scot_deaths %>%
  filter(dow %in% c("Monday", "Tuesday"))

non_hol_perc <- scot_deaths_montues %>% 
  filter(bank_hol == "No") %>%
  group_by(dow) %>%
  summarise(new_deaths_tot = sum(new_deaths)) %>%
  mutate(perc = new_deaths_tot/sum(new_deaths_tot))

non_hol_perc_deaths_hps <- non_hol_perc
  
scot_deaths_adj <- scot_deaths_montues %>% 
  filter(bank_hol == "Yes") %>%
  mutate(week_no = week(date)) %>%
  group_by(week_no) %>%
  mutate(new_deaths_tot = sum(new_deaths)) %>%
  mutate(new_deaths_adj = case_when(dow == "Monday" ~ new_deaths_tot*non_hol_perc$perc[1], 
                                   dow == "Tuesday" ~ new_deaths_tot*non_hol_perc$perc[2])) %>%
  ungroup() %>%
  select(date, new_deaths_adj)

scot_deaths <- left_join(scot_deaths, scot_deaths_adj, by = "date") %>%
  mutate(new_deaths_adj = coalesce(new_deaths_adj,new_deaths)) 

# Scottish tests
scot_tests <- read_excel(path, sheet = "CPT & DPC") 
#scot_tests$Cases[nrow(scot_tests)] <- 4565 #REMOVE

scot_tests  <- scot_tests %>%
  rename("Conducted" = Tests, 
         "Total Positive" = Cases,
         "deaths_per_case" = DPC,
         "cases_per_test" = CPT) %>%
  mutate("Conducted today" = Conducted - replace_na(lag(Conducted), 0)) %>%
  mutate("Total Negative" = Conducted - `Total Positive`) %>%
  mutate("Positive" = `Total Positive` - replace_na(lag(`Total Positive`), 0),
         "Negative" = `Total Negative` - replace_na(lag(`Total Negative`), 0)) 
scot_tests$Positive[scot_tests$Date == ymd("2020-03-12")] <- 24
scot_tests$Negative[scot_tests$Date == ymd("2020-03-12")] <- 552

  
scot_tests_long <- scot_tests %>%
pivot_longer(cols = Positive:Negative, names_to = "Result", values_to = "Number") %>%
mutate(Result = factor(Result, levels = c("Positive", "Negative")))


# Plot colours
#nice_blue <- palette_OkabeIto[5]
#nice_red <- palette_OkabeIto[6]  

nice_blue <- "#0072B2"
nice_red <- "#D55E00"
nice_green <- "#009E73"
nice_orange <- "#E69F00"

#The will affect ratios for 13/20/27 April, 4/11/18/25 May and 1/8 June. 
#public_hols_keep <- unique(c(c(bank_hol+days(6),
#                    bank_hol+days(7)), dmy(c("12/4/2020","13/4/2020","14/4/2020","26/4/2020", #"27/4/2020","28/4/2020", "3/5/2020","4/5/2020","5/5/2020","17/5/2020", "18/5/2020", #"19/5/2020"))))
public_hols_keep_ci <- dmy(c("13/4/2020", "20/4/2020", "27/4/2020","4/5/2020", "11/05/2020", "18/5/2020", "25/5/2020", "1/6/2020", "8/6/2020"))
public_hols_keep <- c(public_hols_keep_ci, public_hols_keep_ci+days(1), public_hols_keep_ci-days(1)) %>% sort()
public_hols_keep_line <- c(public_hols_keep_ci, public_hols_keep_ci-days(1)) %>% sort()

# Weekly Growth 
wk_gr_newcases <- scot_data %>% select(new_cases, date) %>%
      mutate(outcome = new_cases) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
       mutate(ratio_under_1 = ratio_m<1) %>%
       mutate(type = "raw")

wk_gr_newcases_adj <- scot_data %>% select(new_cases_adj, date) %>%
      mutate(outcome = new_cases_adj) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
       mutate(ratio_under_1 = ratio_m<1) %>%
       mutate(type = "adjusted") %>%
       mutate(ratio_m = case_when(date %in% public_hols_keep ~ ratio_m),
              lci = case_when(date %in% public_hols_keep_ci ~ lci),
              uci = case_when(date %in% public_hols_keep_ci ~ uci)) 


wk_gr_newcases_adj <- bind_rows(wk_gr_newcases, wk_gr_newcases_adj) %>% mutate(source = "HPS")

#table(filter(wk_gr_newcases, date >= max(ymd(wk_gr_newcases$date))-20)$significance)
#table(filter(wk_gr_newcases, date >= max(ymd(wk_gr_newcases$date))-20)$ratio_under_1)

wk_gr_newdeaths <- scot_deaths %>% select(new_deaths, date) %>%
      mutate(outcome = new_deaths) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
  mutate(ratio_under_1 = ratio_m<1) %>%
       mutate(type = "raw")

wk_gr_newdeaths_adj <- scot_deaths %>% select(new_deaths_adj, date) %>%
      mutate(outcome = new_deaths_adj) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
  mutate(ratio_under_1 = ratio_m<1) %>%
       mutate(type = "adjusted") %>%
       mutate(ratio_m = case_when(date %in% public_hols_keep ~ ratio_m),
              lci = case_when(date %in% public_hols_keep_ci ~ lci),
              uci = case_when(date %in% public_hols_keep_ci ~ uci))

wk_gr_newdeaths_adj_mainplot <- scot_deaths %>% select(new_deaths_adj, date) %>%
      mutate(outcome = new_deaths_adj) %>%
       weekly_ratios() %>%
      mutate(ci = case_when(date %in% public_hols_keep_ci ~ "Adjusted",
                            !date %in% public_hols_keep_ci ~ "Unadjusted"),
             line =case_when(date %in% public_hols_keep_line ~ "Adjusted",
                            !date %in% public_hols_keep_line ~ "Unadjusted") )

wk_gr_newdeaths_adj_adj <- wk_gr_newdeaths_adj %>%
               mutate(adjusted = date %in% public_hols_keep_ci,
                      adjusted_line = date %in% public_hols_keep)

wk_gr_newdeaths_adj <- bind_rows(wk_gr_newdeaths, wk_gr_newdeaths_adj) %>% mutate(source = "HPS")

#table(filter(wk_gr_newdeaths, date >= max(ymd(wk_gr_newdeaths$date))-20)$significance)
#table(filter(wk_gr_newdeaths, date >= max(ymd(wk_gr_newdeaths$date))-20)$ratio_under_1)


wk_gr_newdeaths_nrs <- nrs_deaths_covid_raw_daily %>% select(new_deaths, date) %>%
      mutate(outcome = new_deaths) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
       mutate(type = "raw")

wk_gr_newdeaths_nrs_adj <- nrs_deaths_covid_raw_daily %>% select(new_deaths_adj, date) %>%
      mutate(outcome = new_deaths_adj) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
       mutate(type = "adjusted") %>%
       mutate(ratio_m = case_when(date %in% public_hols_keep ~ ratio_m),
              lci = case_when(date %in% public_hols_keep_ci ~ lci),
              uci = case_when(date %in% public_hols_keep_ci ~ uci))

wk_gr_newdeaths__nrs_adj_adj <- wk_gr_newdeaths_nrs_adj %>%
               mutate(adjusted = date %in% public_hols_keep_ci)

wk_gr_newdeaths_nrs_adj<- bind_rows(wk_gr_newdeaths_nrs, wk_gr_newdeaths_nrs_adj) %>% mutate(source = "NRS")

wk_gr_newdeaths_nrs_hps <- bind_rows(wk_gr_newdeaths %>% mutate(source = "HPS"),
wk_gr_newdeaths_nrs %>% mutate(source = "NRS"))

wk_gr_newdeaths_nrs_hps_adj <- bind_rows(wk_gr_newdeaths_adj,
wk_gr_newdeaths_nrs_adj)

#nrs by age
nrs_deaths_covid_week_age <- nrs_deaths_covid_raw %>% 
                             slice(9:15) %>%
                             pivot_longer(cols = `43829`:names(nrs_deaths_covid_raw)[ncol(nrs_deaths_covid_raw)-1],
               names_to = "week_beginning", 
               values_to = "weekly_deaths_total") %>%
  #mutate(Details = case_when(`Week beginning` == "Deaths involving COVID-194" ~ "All",
  #                           TRUE ~ Details)) %>%
  mutate(week_beginning = janitor::excel_numeric_to_date(parse_number(week_beginning))) %>%
  select(-`Week beginning`) %>%
  rename("Age Group" = Details) %>%
  #filter(Location != "Other institution") %>%
  select(`Age Group`, week_beginning, weekly_deaths_total) %>%
  mutate(week_ending = week_beginning + 6)

wk_gr_nrs_deaths_age <- nrs_deaths_covid_week_age %>%
  group_by(`Age Group`) %>%
  mutate(week_ending_prev = lag(weekly_deaths_total, 1)) %>%
  filter(!is.na(week_ending_prev) & week_ending_prev >0) %>% 
  ungroup() %>%
  group_by(`Age Group`, week_ending) %>%
  nest() %>%
  mutate(tab = map(data, ~ weekly_ratios_ci(.x$weekly_deaths_total, .x$week_ending_prev))) %>%
  mutate(ratio_m = map_dbl(tab, ~parse_number(as.character((.x[1,2]))))) %>% 
  mutate(lci = map_dbl(tab, ~parse_number(as.character((.x[2,2]))))) %>% 
  mutate(uci = map_dbl(tab, ~parse_number(as.character((.x[3,2]))))) %>% 
  ungroup() %>%
  select(`Age Group`, week_ending, ratio_m, lci, uci) %>%
  mutate(comparison = if_else(ratio_m >1, "greater than", "less than")) %>%
  significance_wk() %>%
  rename(date = week_ending)


#nrs by location
nrs_deaths_covid_week <- nrs_deaths_covid_raw %>% 
  filter(`Week beginning` == "Deaths involving COVID-194" | 
          Details %in% c("Care Home", "Home / Non-institution", "Hospital", "Other institution")) %>%
  pivot_longer(cols = `43829`:names(nrs_deaths_covid_raw)[ncol(nrs_deaths_covid_raw)-1],
               names_to = "week_beginning", 
               values_to = "weekly_deaths_total") %>%
  mutate(Details = case_when(`Week beginning` == "Deaths involving COVID-194" ~ "All",
                             TRUE ~ Details)) %>%
  mutate(week_beginning = janitor::excel_numeric_to_date(parse_number(week_beginning))) %>%
  mutate(Type = "Covid19 deaths") %>%
  rename("Location" = Details) %>%
  filter(Location != "Other institution") %>%
  select(Location, week_beginning, weekly_deaths_total, Type) %>%
  mutate(week_ending = week_beginning + 6)

wk_gr_nrs_deaths <- nrs_deaths_covid_week %>%
  group_by(Location) %>%
  mutate(week_ending_prev = lag(weekly_deaths_total, 1)) %>%
  filter(!is.na(week_ending_prev) & week_ending_prev >0) %>% 
  #pivot_longer(c(weekly_deaths_total,week_ending_prev), names_to = "week", values_to = "new_numbers_week") %>%
  #mutate(week = factor(week, levels = c("week_ending_prev", "weekly_deaths_total"))) %>%
  ungroup() %>%
  group_by(Location, week_ending) %>%
  nest() %>%
  mutate(tab = map(data, ~ weekly_ratios_ci(.x$weekly_deaths_total, .x$week_ending_prev))) %>%
  mutate(ratio_m = map_dbl(tab, ~parse_number(as.character((.x[1,2]))))) %>% 
  mutate(lci = map_dbl(tab, ~parse_number(as.character((.x[2,2]))))) %>% 
  mutate(uci = map_dbl(tab, ~parse_number(as.character((.x[3,2]))))) %>% 
  ungroup() %>%
  select(Location, week_ending, ratio_m, lci, uci) %>%
  mutate(comparison = if_else(ratio_m >1, "greater than", "less than")) %>%
  significance_wk() %>%
  rename(date = week_ending)

wk_gr_nrs_deaths <- left_join(wk_gr_nrs_deaths,
  nrs_deaths_covid_week %>% 
  filter(weekly_deaths_total >0) %>%
  rename("change" = weekly_deaths_total, 
         "date" = "week_ending") %>%
  select(change, date, Location))

wk_gr_newdeaths_latest <- wk_gr_newdeaths %>% mutate(source = "HPS", Location = "HPS") %>% filter(date == max(wk_gr_newdeaths$date))

wk_gr_nps_nrs <- bind_rows(wk_gr_newdeaths %>% mutate(source = "HPS")  %>%
                           filter(date %in% wk_gr_nrs_deaths$date), 
                           wk_gr_nrs_deaths %>% mutate(source = "NRS"))
wk_gr_nps_nrs <- wk_gr_nps_nrs %>%
                 mutate(Location = case_when( source == "HPS" ~ source, 
                                              Location == "All" ~ "NRS",
                                             TRUE ~ Location))
  
```

```{r}
## Comparison numbers
# Cases
cases_comp <- filter(wk_gr_newcases, date %in% c(max(wk_gr_newcases$date), max(wk_gr_newcases$date)-21)) %>%
  mutate(dow = weekdays(date))

dcl_inc <- ifelse(cases_comp$change[2] - cases_comp$change[1] > 0, "increased", "declined")
fall_inc <- ifelse(cases_comp$change[2] - cases_comp$change[1] > 0, "an increase", "a fall")
fall_inc_no <- 100*abs(cases_comp$change[2] - cases_comp$change[1])/cases_comp$change[1]
fall_inc_no <- formatC(round(fall_inc_no,2), format='f', digits=1)

# Deaths HPS
deaths_hps_comp <- filter(wk_gr_newdeaths, date %in% c(max(wk_gr_newdeaths$date), max(wk_gr_newdeaths$date)-21)) %>%
  mutate(dow = weekdays(date))

dcl_inc_hps <- ifelse(deaths_hps_comp$change[2] - deaths_hps_comp$change[1] > 0, "increased", "declined")
fall_inc_hps <- ifelse(deaths_hps_comp$change[2] - deaths_hps_comp$change[1] > 0, "an increase", "a fall")
fall_inc_no_hps <- 100*abs(deaths_hps_comp$change[2] - deaths_hps_comp$change[1])/deaths_hps_comp$change[1]
fall_inc_no_hps <- formatC(round(fall_inc_no_hps,2), format='f', digits=1)

# Deaths NRS
deaths_nrs_comp <- filter(wk_gr_newdeaths_nrs, date %in% c(max(wk_gr_newdeaths_nrs$date), max(wk_gr_newdeaths_nrs$date)-21)) %>%
  mutate(dow = weekdays(date))

dcl_inc_nrs <- ifelse(deaths_nrs_comp$change[2] - deaths_nrs_comp$change[1] > 0, "increased", "declined")
fall_inc_nrs <- ifelse(deaths_nrs_comp$change[2] - deaths_nrs_comp$change[1] > 0, "an increase", "a fall")
fall_inc_no_nrs <- 100*abs(deaths_nrs_comp$change[2] - deaths_nrs_comp$change[1])/deaths_nrs_comp$change[1]
fall_inc_no_nrs <- formatC(round(fall_inc_no_nrs,2), format='f', digits=1)

# Hosp hosp icu
hosp_comp <- filter(data_hosp, date %in% c(max(data_hosp$date), max(data_hosp$date)-21)) %>%
  mutate(dow = weekdays(date))

dcl_inc_hosp <- ifelse(hosp_comp$Hospital_confirmed[2] - hosp_comp$Hospital_confirmed[1] > 0, "increased", "declined")
fall_inc_hosp <- ifelse(hosp_comp$Hospital_confirmed[2] - hosp_comp$Hospital_confirmed[1] > 0, "an increase", "a fall")
fall_inc_no_hosp <- 100*abs(hosp_comp$Hospital_confirmed[2] - hosp_comp$Hospital_confirmed[1])/hosp_comp$Hospital_confirmed[1]
fall_inc_no_hosp <- formatC(round(fall_inc_no_hosp,2), format='f', digits=1)

dcl_inc_icu <- ifelse(hosp_comp$ICU_total[2] - hosp_comp$ICU_total[1] > 0, "increased", "declined")
fall_inc_icu <- ifelse(hosp_comp$ICU_total[2] - hosp_comp$ICU_total[1] > 0, "an increase", "a fall")
fall_inc_no_icu <- 100*abs(hosp_comp$ICU_total[2] - hosp_comp$ICU_total[1])/hosp_comp$ICU_total[1]
fall_inc_no_icu <- formatC(round(fall_inc_no_icu,2), format='f', digits=1)
```


The number of cases over the past 7 days has `r dcl_inc` to `r cases_comp$change[2]` today from `r cases_comp$change[1]` three weeks previously, `r fall_inc` of `r fall_inc_no`%.

The number of deaths recorded by HPS over the past 7 days has `r dcl_inc_hps` to `r deaths_hps_comp$change[2]` today from `r deaths_hps_comp$change[1]` three weeks previously, `r fall_inc_hps` of `r fall_inc_no_hps`%.

The number of deaths recorded by NRS over the past 7 days has `r dcl_inc_nrs` to `r deaths_nrs_comp$change[2]` today from `r deaths_nrs_comp$change[1]` three weeks previously, `r fall_inc_nrs` of `r fall_inc_no_nrs`%.

The number of patients in hospital with confirmed COVID-19 on `r format(max(data_hosp$date), '%d/%m/%Y')` has `r dcl_inc_hosp` to `r hosp_comp$Hospital_confirmed[2]` today from `r hosp_comp$Hospital_confirmed[1]` three weeks previously, `r fall_inc_hosp` of `r fall_inc_no_hosp`%.

The number of patients in ICU with confirmed or suspect COVID-19 on `r format(max(data_hosp$date), '%d/%m/%Y')` has `r dcl_inc_icu` to `r hosp_comp$ICU_total[2]` today from `r hosp_comp$ICU_total[1]` three weeks previously, `r fall_inc_icu` of `r fall_inc_no_icu`%.


### Over the last 21 days...

### Cases
```{r}
table(filter(wk_gr_newcases, date >= max(ymd(wk_gr_newcases$date))-20)$ratio_under_1) %>% kable(caption = "Ratio under 1")
table(filter(wk_gr_newcases, date >= max(ymd(wk_gr_newcases$date))-20)$significance) %>% kable(caption = "Significance")
```

### Deaths HPS
```{r}
table(filter(wk_gr_newdeaths, date >= max(ymd(wk_gr_newdeaths$date))-20)$ratio_under_1) %>% kable(caption = "Ratio under 1")
table(filter(wk_gr_newdeaths, date >= max(ymd(wk_gr_newdeaths$date))-20)$significance) %>% kable(caption = "Significance")
```

### Hospital
```{r}
table(filter(wk_gr_hosp_hosp, date >= max(ymd(wk_gr_hosp_hosp$date))-20)$ratio_under_1) %>% kable(caption = "Ratio under 1")
table(filter(wk_gr_hosp_hosp, date >= max(ymd(wk_gr_hosp_hosp$date))-20)$significance)  %>% kable(caption = "Significance")
```

### ICU
```{r}
table(filter(wk_gr_hosp_icu, date >= max(ymd(wk_gr_hosp_icu$date))-20)$ratio_under_1) %>% kable(caption = "Ratio under 1")
table(filter(wk_gr_hosp_icu, date >= max(ymd(wk_gr_hosp_icu$date))-20)$significance)  %>% kable(caption = "Significance")
```

\newpage 
### Other figures
  
```{r, eval=FALSE}
nrs_now<- filter(nrs_deaths_covid, week_beginning == max(nrs_deaths_covid$week_beginning))$weekly_deaths_total
nrs_ldwk1 <- filter(nrs_deaths_covid, week_beginning == "2020-03-23")$weekly_deaths_total

nrs_deaths_covid %>%
  filter(week_beginning %in% 
c(ymd("2020-03-23"), max(nrs_deaths_covid_week_age$week_beginning))) %>% 
pivot_wider(names_from = week_beginning, values_from = weekly_deaths_total) %>%
mutate("Perc diff" = round(100*(`2020-05-18`-`2020-03-23`)/`2020-03-23`,2)) %>%
  kable(caption = "New deaths overall (Source: NRS)")

nrs_deaths_covid_week_age %>% 
filter(week_beginning %in% 
c(ymd("2020-03-23"), max(nrs_deaths_covid_week_age$week_beginning))) %>%
pivot_wider(names_from = week_beginning, values_from = weekly_deaths_total, id_cols = `Age Group`) %>%
mutate("Perc diff" = round(100*(`2020-05-18`-`2020-03-23`)/`2020-03-23`,2)) %>%
  kable(caption = "New deaths by Age group (Source: NRS)")

nrs_deaths_covid_week %>%
  filter(week_beginning %in% 
c(ymd("2020-03-23"), max(nrs_deaths_covid_week_age$week_beginning)) & Location %in% c("All", "Care Home")) %>%
  pivot_wider(names_from = Location, values_from = weekly_deaths_total, id_cols = week_beginning) %>%
  mutate("Non care home" = All - `Care Home`) %>%
  select(-All) %>%
  pivot_longer(`Care Home`:`Non care home`, names_to = "Location", values_to = "weekly_deaths_total") %>%
  pivot_wider(names_from = week_beginning, values_from = weekly_deaths_total, id_cols = Location) %>%
mutate("Perc diff" = round(100*(`2020-05-18`-`2020-03-23`)/`2020-03-23`,2)) %>%
  kable(caption = "New deaths - Care Homes Vs Non-care homes (Source: NRS)")

scot_data %>%
  mutate(week_beginning = floor_date(date, unit="week", week_start = 1)) %>%
  group_by(week_beginning) %>%
  summarise(weekly_cases_total = sum(new_cases)) %>% 
  filter(week_beginning %in% 
c(ymd("2020-03-23"), max(nrs_deaths_covid_week_age$week_beginning))) %>%
  pivot_wider(names_from = week_beginning, values_from = weekly_cases_total) %>%
mutate("Perc diff" = round(100*(`2020-05-18`-`2020-03-23`)/`2020-03-23`,2)) %>%
  kable(caption = "New cases (Source: HPS)")
```


\newpage

\thispagestyle{empty}

```{r casesdeathsicuratios, fig.width=5.5, fig.height= 9, fig.cap="Ratios of weekly new cases, deaths, ICU occupancy and hospitalisations vs respective counts for previous week; graphics prepared by The University of Edinburgh"}
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New Cases (Source: HPS)") +
  theme_bw(base_size = 10) +
  theme(plot.margin = unit(c(0.2,0.5,0.2,0.3), "cm"))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-20)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "8 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.57, y = .54, width = .37, height = .3)

fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 10) + 
  theme(plot.margin = unit(c(0.2,0.5,0.2,0.2), "cm"))

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-20)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "8 days", labels = date_format("%b %d"),expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.57, y = .54, width = .37, height = .3)


fig3 <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Patients in ICU (Source: NHS Boards)") +
  theme_bw(base_size = 10)  + 
  theme(plot.margin = unit(c(0.2,0.4,0.2,0.3), "cm"))



fig3_inset <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_icu$date)-20)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_x_date(date_breaks = "8 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig3 <- ggdraw() +
  draw_plot(fig3) +
  draw_plot(fig3_inset, x = 0.57, y = .54, width = .37, height = .3)

fig4 <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.1) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Patients in Hospital (Source: NHS Boards)") +
  theme_bw(base_size = 10)  + 
  theme(plot.margin = unit(c(0.2,0.4,0.2,0.3), "cm"))

fig4_inset <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_hosp$date)-20)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "8 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig4 <- ggdraw() +
  draw_plot(fig4) +
  draw_plot(fig4_inset, x = 0.57, y = .54, width = .37, height = .3)

gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 4) 

```

```{r, fig.width=15, fig.height=9, fig.cap="Ratios of new cases, new deaths, ICU occupancy and hospitalisations vs respective counts for previous time period. Cases and deaths are comparisons of weekly sums; ICU occupancy and hospitalisations are week over week comparisons of daily counts. Graphics provided by The University of Edinburgh."}
gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 2, ncol = 2) 
```

```{r, fig.width=15, fig.height=9, fig.cap="Ratios of new cases, new deaths, ICU occupancy and hospitalisations vs respective counts for previous time period. Cases and deaths are comparisons of weekly sums; ICU occupancy and hospitalisations are week over week comparisons of daily counts. Graphics provided by The University of Edinburgh.", eval=FALSE}
png("weekly_ratios_landscape.png", width = 11.69, height = 8.27, units = "in", res = 400)
#gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 2, ncol = 2, bottom = "Ratios of new cases, new deaths, hospitalisations and ICU occupancy vs respective counts for previous time period. Cases and deaths are comparisons \nof weekly sums; Hospitalisations and ICU occupancy are week over week comparisons of daily counts. Graphics provided by The University of Edinburgh.")
gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 2, ncol = 2, bottom = "Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 scale.  Day-of-the-week ratios for hospitalisations and \nICU  occupancy are ratios of daily counts, measured 7 days apart and plotted on a log10 scale. Graphics provided by The University of Edinburgh.")
dev.off()
```


```{r}
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  scale_y_continuous(trans = "log10",
                         breaks = c(0.5, 1,3,6),
                         #labels = c(0.5, 1,3,6),
                         expand = expansion(mult = 0.1), limits = c(NA,7)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New Cases (Source: HPS)") +
  theme_bw(base_size = 10) +
  theme(plot.margin = unit(c(0.2,0.5,0.2,0.3), "cm"))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-20)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.51, y = .55, width = .44, height = .29)


fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                     breaks = c(0.5, 1,3,10,30),
                     expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 10) + 
  theme(plot.margin = unit(c(0.2,0.5,0.2,0.2), "cm"))

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-20)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.51, y = .54, width = .44, height = .3)


fig3 <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.1) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                     breaks = c(0.8, 1,2, 3),
                     #labels = c(0.8, 1,2, 3),
                     expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Patients in Hospital (Source: NHS Boards)") +
  theme_bw(base_size = 10)  + 
  theme(plot.margin = unit(c(0.2,0.4,0.2,0.3), "cm"))

fig3_inset <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_hosp$date)-20)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig3 <- ggdraw() +
  draw_plot(fig3) +
  draw_plot(fig3_inset, x = 0.51, y = .54, width = .44, height = .3)

fig4 <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_y_continuous(trans = "log10",
                         breaks = c(0.5, 1, 3, 10),
                         #labels = c(0.5, 1, 3, 10),
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Patients in ICU (Source: NHS Boards)") +
  theme_bw(base_size = 10)  + 
  theme(plot.margin = unit(c(0.2,0.4,0.2,0.3), "cm"))



fig4_inset <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_icu$date)-20)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig4 <- ggdraw() +
  draw_plot(fig4) +
  draw_plot(fig4_inset, x = 0.51, y = .54, width = .44, height = .3)


gridExtra::grid.arrange(fig1, fig2, fig3, fig4, nrow = 4) 
```

```{r}
png("weekly_ratios_portrait_4email.png", height = 11.69, width = 7, units = "in", res = 400)
gridExtra::grid.arrange(fig1, fig2, fig3, fig4, nrow = 4, bottom = "Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 \nscale. Day-of-the-week ratios for number of patients in hospital and ICU are ratios of daily \ncounts, measured 7 days apart and plotted on a log10 scale. Figure insets highlight ratios \nand trends for the last 21 days. Graphics provided by The University of Edinburgh.")
dev.off()
```

```{r}
png("weekly_ratios_landscape_4email.png", width = 11.69, height = 7, units = "in", res = 400)
gridExtra::grid.arrange(fig1, fig2, fig3, fig4, nrow = 2, ncol = 2, bottom = "Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 \nscale. Day-of-the-week ratios for number of patients in hospital and ICU are ratios of daily \ncounts, measured 7 days apart and plotted on a log10 scale. Figure insets highlight ratios \nand trends for the last 21 days. Graphics provided by The University of Edinburgh.")
dev.off()
```



```{r fig.width=5.5, fig.height= 9, fig.cap="Ratios of new cases, new deaths, ICU occupancy and hospitalisations vs respective counts for previous time period. Cases and deaths are comparisons of weekly sums; ICU occupancy and hospitalisations are week over week comparisons of daily counts. Graphics provided by The University of Edinburgh."}
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  scale_y_continuous(trans = "log10",
                         breaks = c(0.5, 1,3,6),
                         #labels = c(0.5, 1,3,6),
                         expand = expansion(mult = 0.1), limits = c(NA,7)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New Cases (Source: HPS)") +
  theme_bw(base_size = 10) +
  theme(plot.margin = unit(c(0.2,0.5,0.2,0.3), "cm"))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-20)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.51, y = .55, width = .44, height = .29)


fig2 <- ggplot(wk_gr_newdeaths_adj_mainplot %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, alpha = 0.4, aes(colour = ci)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(aes(colour = line, group = 1)) +
  geom_point(size = 1, aes(colour = ci)) +
  scale_y_continuous(trans = "log10",
                     expand = expansion(mult = 0.1),
                     breaks = c(0.5, 1,3,10,30),
                     #labels = c(0.5, 1,3,10,30)
                     ) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  scale_colour_manual(values = c("grey50", nice_red)) +
  scale_fill_manual(values = c("grey50", nice_red)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 10) + 
  theme(plot.margin = unit(c(0.2,0.5,0.2,0.2), "cm"), 
        legend.position = "none")

fig2_inset <- ggplot(wk_gr_newdeaths_adj_mainplot %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= max(wk_gr_newdeaths_adj$date)-20)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, alpha = 0.4, aes(colour = ci)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(aes(colour = line, group = 1)) +
  geom_point(size = 1, aes(colour = ci)) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")#,expand = expansion(mult = 0.1)
               ) +
  scale_colour_manual(values = c("grey50", nice_red)) +
  scale_fill_manual(values = c("grey50", nice_red)) +
    scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"), 
        legend.position = "none")

# fig2 <- ggplot(wk_gr_newdeaths_adj %>% filter(!is.na(ratio) & 
#               !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
#   aes(x = date, y = ratio_m,
#     ymin = lci, ymax = uci, colour = type, fill = type) +
#   geom_errorbar(width = 0, alpha = 0.4) +
#   geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
#   geom_line() +
#   geom_point(size = 1) +
#   scale_y_continuous(trans = "log10",
#   scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
#                limits = c(dmy("23-03-2020"), NA)) +
#   scale_colour_manual(values = c("grey50", nice_red)) +
#   scale_fill_manual(values = c("grey50", nice_red)) +
#   labs( x = "End Date of Current Week",
#         y = "Weekly Ratio",
#         subtitle = "New Deaths (Source: HPS)") +
#   theme_bw(base_size = 10) + 
#   theme(plot.margin = unit(c(0.2,0.5,0.2,0.2), "cm"), 
#         legend.position = "none")
# 
# fig2_inset <- ggplot(wk_gr_newdeaths_adj %>% filter(!is.na(ratio) & 
#               !is.infinite(ratio) & date >= max(wk_gr_newdeaths_adj$date)-20)) +
#   aes(x = date, y = ratio_m,
#     ymin = lci, ymax = uci, colour = type, fill = type) +
#   geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=0.3)) +
#   geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
#   geom_line(position=position_dodge(width=0.3)) +
#   geom_point(size = 1, position=position_dodge(width=0.3)) +
#   scale_x_date(date_breaks = "8 days", labels = date_format("%b %d")#,expand = expansion(mult = 0.1)
#                ) +
#   scale_colour_manual(values = c("grey50", nice_red)) +
#   scale_fill_manual(values = c("grey50", nice_red)) +
#     scale_y_continuous(trans = "log10") +
#   labs( x = "",
#         y = "") +
#   theme_bw(base_size = 8) +
#   theme(plot.margin = margin(0, 0, -1, 0, "cm"), 
#         legend.position = "none")

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.51, y = .54, width = .44, height = .3)


fig3 <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.1) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                     breaks = c(0.8, 1,2, 3),
                     #labels = c(0.8, 1,2, 3),
                     expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Patients in Hospital (Source: NHS Boards)") +
  theme_bw(base_size = 10)  + 
  theme(plot.margin = unit(c(0.2,0.4,0.2,0.3), "cm"))

fig3_inset <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_hosp$date)-20)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig3 <- ggdraw() +
  draw_plot(fig3) +
  draw_plot(fig3_inset, x = 0.51, y = .54, width = .44, height = .3)

fig4 <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_y_continuous(trans = "log10",
                         breaks = c(0.5, 1, 3, 10),
                         #labels = c(0.5, 1, 3, 10),
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d"),
               limits = c(dmy("23-03-2020"), NA)) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Patients in ICU (Source: NHS Boards)") +
  theme_bw(base_size = 10)  + 
  theme(plot.margin = unit(c(0.2,0.4,0.2,0.3), "cm"))



fig4_inset <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_icu$date)-20)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_x_date(date_breaks = "9 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig4 <- ggdraw() +
  draw_plot(fig4) +
  draw_plot(fig4_inset, x = 0.51, y = .54, width = .44, height = .3)


gridExtra::grid.arrange(fig1, fig2, fig3, fig4, nrow = 4) 
```

```{r}
library(grid)
png("weekly_ratios_portrait_4email_adj.png", height = 11.69, width = 7, units = "in", res = 400)
gridExtra::grid.arrange(fig1, fig2, fig3, fig4, nrow = 4, bottom = textGrob("Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 scale. \nDay-of-the-week ratios for number of patients in hospital and ICU are ratios of daily counts, measured \n7 days apart and plotted on a log10 scale. Entries in grey are adjusted for public holiday effects, see \nCaveats. Figure insets highlight ratios and trends for the last 21 days. Graphics provided by \nThe University of Edinburgh.",gp=gpar(fontsize=11,font=3)))
dev.off()
```

```{r}
library(grid)
png("weekly_ratios_landscape_4email_adj.png", width = 11.69, height = 7, units = "in", res = 400)
gridExtra::grid.arrange(fig1, fig2, fig3, fig4, nrow = 2, ncol = 2, bottom = textGrob("Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 scale. Day-of-the-week ratios \nfor number of patients in hospital and ICU are ratios of daily counts, measured 7 days apart and plotted on a log10 scale. \nEntries in grey are adjusted for public holiday effects, see Caveats. Figure insets highlight ratios and trends for the last 21 days. \nGraphics provided by The University of Edinburgh.",gp=gpar(fontsize=11,font=3)))
dev.off()
```




```{r fig.width=5.5, fig.height= 9, fig.cap="Ratios of new cases, new deaths, ICU occupancy and hospitalisations vs respective counts for previous time period. Cases and deaths are comparisons of weekly sums; ICU occupancy and hospitalisations are week over week comparisons of daily counts. Graphics provided by The University of Edinburgh.", eval=FALSE}
adjusted <- read_excel("Adjusted ratios.xlsx", skip = 1) %>%
  select(Date, `Ratio...13`, `Lower...14`, `Upper...15`) %>%
  rename(ratio_m = `Ratio...13`, 
         lci = `Lower...14`, 
         uci = `Upper...15`) %>%
  mutate(date = ymd(Date)) %>%
  select(-Date) %>%
  filter(date %in% dmy(c("04/05/2020", "11/05/2020"))) %>%
  bind_rows(wk_gr_newdeaths %>% select(date, ratio_m) %>%filter(date %in% dmy(c("03/05/2020","05/05/2020", "10/05/2020", "12/05/2020"))))
#adjusted$date <-  ymd_h(paste(adjusted$date, "0")) + hours(4)

adjusted_1 <- filter(adjusted, date %in% dmy(c("03/05/2020","04/05/2020","05/05/2020"))) %>%
  mutate(type = "adjusted_1")
adjusted_2 <- filter(adjusted, date %in% dmy(c("10/05/2020","11/05/2020","12/05/2020"))) %>%
  mutate(type = "adjusted_2")
wk_gr_newdeaths <- wk_gr_newdeaths %>%
  mutate(type = "unadjusted")

wk_gr_newdeaths_adj <- bind_rows(wk_gr_newdeaths, adjusted_1, adjusted_2)

tiff("weekly_ratios_newdeaths_adjusted.tiff", height = 11.69/4, width = 5, units = "in", res = 400)
fig2 <- ggplot(wk_gr_newdeaths_adj %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci, colour = type, fill = type) +
  geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=0.3)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(position=position_dodge(width=0.3)) +
  geom_point(size = 0.5, position=position_dodge(width=0.3)) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  scale_colour_manual(values = c("grey50", "grey51", "#D55E00")) +
  scale_fill_manual(values = c("grey50", "grey51", "#D55E00")) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 8) +
  theme(legend.position = "none")

fig2_inset <- ggplot(wk_gr_newdeaths_adj %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= max(wk_gr_newdeaths$date)-6)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci, colour = type, fill = type) +
  geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=0.2)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(position=position_dodge(width=0.2)) +
  geom_point(size = 1, position=position_dodge(width=0.2)) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"),expand = expansion(mult = 0.1)) +
  scale_colour_manual(values = c("grey51", "#D55E00")) +
  scale_fill_manual(values = c("grey51", "#D55E00")) +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"), 
        legend.position = "none")

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.57, y = .52, width = .40, height = .3)
fig2
dev.off()

tiff("weekly_ratios_newcases.tiff", height = 11.69/4, width = 5, units = "in", res = 400)
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New Cases (Source: HPS)") +
  theme_bw(base_size = 10) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 14, b = 0, l = 0)))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.57, y = .52, width = .40, height = .3)

fig1
dev.off()


tiff("weekly_ratios_newdeaths.tiff", height = 11.69/4, width = 5, units = "in", res = 400)
fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 10)

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"),expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.57, y = .52, width = .40, height = .3)
fig2
dev.off()

tiff("dayofweek_ratios_icuoccup.tiff", height = 11.69/4, width = 5, units = "in", res = 400)
fig3 <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "ICU Occupancy (Source: NHS Boards)") +
  theme_bw(base_size = 10) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 7, b = 0, l = 0)))



fig3_inset <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_icu$date)-8)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig3 <- ggdraw() +
  draw_plot(fig3) +
  draw_plot(fig3_inset, x = 0.57, y = .52, width = .40, height = .3)
fig3
dev.off()

tiff("dayofweek_ratios_hospoccup.tiff", height = 11.69/4, width = 5, units = "in", res = 400)

fig4 <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Hospital Occupancy (Source: NHS Boards)") +
  theme_bw(base_size = 10) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 14, b = 0, l = 0)))


fig4_inset <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_hosp$date)-8)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 8) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig4 <- ggdraw() +
  draw_plot(fig4) +
  draw_plot(fig4_inset, x = 0.57, y = 0.52, width = 0.40, height = 0.3)

fig4
dev.off()
```

```{r, eval = FALSE}
png("weekly_ratios_portrait.png", height = 11.69, width = 7, units = "in", res = 400)
#gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 4, bottom = "Ratios of new cases, new deaths, hospitalisations and ICU occupancy  vs respective \ncounts for previous time period. Cases and deaths are comparisons of weekly sums; \nHospitalisations and ICU occupancy are week over week comparisons of daily counts. \nGraphics provided by The University of Edinburgh.") 
gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 4, bottom = "Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 \nscale. Day-of-the-week ratios for hospitalisations and ICU occupancy are ratios of \ndaily counts, measured 7 days apart and plotted on a log10 scale. Graphics provided by \nThe University of Edinburgh.")
dev.off()
```

```{r, eval = FALSE}
tiff("weekly_ratios_portrait.tiff", height = 11.69, width = 7, units = "in", res = 400)
gridExtra::grid.arrange(fig1, fig2, fig4, fig3, nrow = 4, bottom = "Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 \nscale. Day-of-the-week ratios for hospital occupancy and ICU occupancy are ratios of daily \ncounts, measured 7 days apart and plotted on a log10 scale. Figure insets highlight more \nrecent ratios and trends. Graphics provided by The University of Edinburgh.")
dev.off()
```


