---
title: "Monitoring the COVID-19 Pandemic in Scotland"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
header-includes:
   - \usepackage{booktabs}
   - \usepackage{float}
   - \floatplacement{figure}{H}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: no
---


```{r setup, include=FALSE}
#date: "`r format(Sys.time(), '%d %B, %Y')`" 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.pos = 'H')
# Load packages
library(flexdashboard) ; library(shiny) ; library(readr); library(dplyr); library(tidyr); library(purrr); library(forcats); library(stringr); library(htmlwidgets); library(lubridate); library(sf); library(RcppRoll); library(plotly); library(shinythemes);library(leaflet); library(classInt); library(ggrepel); library(scales); library(leaflet.extras); library(RColorBrewer);
library(colorblindr); library(readxl);library(spatstat.utils);library(httr);library(cowplot)

#source("growth_rate_window.R")
source("weekly_ratios.R")
#source("weekly_ratios_byday.R")


# Import hospital data
## Covid daily hospital data
myurl <- "https://www.gov.scot/binaries/content/documents/govscot/publications/statistics/2020/04/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/documents/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/trends-in-number-of-people-in-hospital-with-confirmed-or-suspected-covid-19/govscot%3Adocument/Data%2BTable%2B%252810-04-2020%2529.xlsx?forceDownload=true"
GET(myurl, write_disk(tmp <- tempfile(fileext = ".xlsx")))

#data_hosp <- read_excel(tmp, sheet = "Table 1", skip = 3)
data_hosp <- read_excel(tmp, sheet = "Table 2 - Hospital Care", skip = 3)

data_hosp <- data_hosp %>%
  rename("date" = "...1",
         "ICU_confirmed" = "Confirmed...2",
         "ICU_suspected" ="Suspected...3", 
         "ICU_total" = "Total...4",
         "Hospital_confirmed" = "Confirmed...5",
         "Hospital_suspected" = "Suspected...6",
         "Hospital_total" = "Total...7" ) %>%
  filter(!is.na(ICU_total)) %>%
  #mutate(date = as.Date(as.numeric(date), origin = "1899-12-30")) %>%
  mutate(date = lubridate::ymd(date)) %>%
  mutate(ICU_confsusp = case_when(is.na(ICU_confirmed) & is.na(ICU_suspected) ~ ICU_total)) %>%
  mutate(Hospital_confsusp = case_when(is.na(Hospital_confirmed) & is.na(Hospital_suspected) ~ Hospital_total)) %>%
  select(date, contains("ICU"), everything()) %>%
  select(date, ICU_total, Hospital_total) 

wk_gr_hosp_icu <- data_hosp %>%
       mutate(ICU_prev = lag(ICU_total , 7), 
         weekdays(date)) %>%
       filter(date >= ymd("2020-03-25")) %>%
       pivot_longer(c(ICU_total, ICU_prev), names_to = "week", values_to = "number") %>%
      group_by(date) %>%
      nest() %>%
      mutate(mod = map(data, ~ glm(number ~ week,
                              data = .x, family = poisson))) %>%
  mutate(ratio_m = map_dbl(mod, ~ exp(coef(.x)["weekICU_total"]))) %>% 
  mutate(lci = map_dbl(mod, ~ exp(confint(.x)["weekICU_total", "2.5 %"]))) %>%
  mutate(uci = map_dbl(mod, ~ exp(confint(.x)["weekICU_total", "97.5 %"]))) %>%
  ungroup() %>%
  select(date, ratio_m, lci, uci) %>%
  mutate(comparison = if_else(ratio_m >1, "greater than", "less than")) %>%
  significance_wk() 


wk_gr_hosp_icu_latest <- filter(wk_gr_hosp_icu, date == max(wk_gr_hosp_icu$date))

wk_gr_hosp_hosp <- data_hosp %>%
       mutate(Hosp_prev = lag(Hospital_total , 7), 
         weekdays(date)) %>%
       filter(date >= ymd("2020-03-25")) %>%
       pivot_longer(c(Hospital_total, Hosp_prev), names_to = "week", values_to = "number") %>%
      group_by(date) %>%
      nest() %>%
      mutate(mod = map(data, ~ glm(number ~ week,
                              data = .x, family = poisson))) %>%
  mutate(ratio_m = map_dbl(mod, ~ exp(coef(.x)["weekHospital_total"]))) %>% 
  mutate(lci = map_dbl(mod, ~ exp(confint(.x)["weekHospital_total", "2.5 %"]))) %>%
  mutate(uci = map_dbl(mod, ~ exp(confint(.x)["weekHospital_total", "97.5 %"]))) %>%
  ungroup() %>%
  select(date, ratio_m, lci, uci) %>%
  mutate(comparison = if_else(ratio_m >1, "greater than", "less than")) %>%
  significance_wk() 


wk_gr_hosp_hosp_latest <- filter(wk_gr_hosp_hosp, date == max(wk_gr_hosp_icu$date))

# Import NRS data
myurl_deaths <- "https://www.nrscotland.gov.uk/files//statistics/covid19/covid-deaths-data-week-17.xlsx"
GET(myurl_deaths, write_disk(tmp <- tempfile(fileext = ".xlsx")))

#NRS weekly
nrs_deaths_covid_raw <- read_excel(tmp, sheet = "Table 1 - COVID deaths", skip = 3) %>%
  select(-`Year to Date`) %>%
  rename("Details" = "...2",
         "Total" = "...21")

nrs_deaths_covid <- nrs_deaths_covid_raw %>% filter(`Week beginning` == "Deaths involving COVID-194") %>%
  pivot_longer(cols = `43829`:names(nrs_deaths_covid_raw)[ncol(nrs_deaths_covid_raw)-1],
               names_to = "week_beginning", 
               values_to = "weekly_deaths_total") %>%
  mutate(week_beginning = janitor::excel_numeric_to_date(parse_number(week_beginning))) %>%
  mutate(Type = "Covid19 deaths") %>%
  select(week_beginning, weekly_deaths_total, Type)

# NRS daily
nrs_deaths_covid_raw_daily <- read_excel(tmp, sheet = "Figure 2 data", skip =2) %>%
  filter(!is.na(Source)) %>%
  mutate(date = janitor::excel_numeric_to_date(parse_number(Date1))) %>%
  filter(Source == "NRS") %>%
  mutate(new_deaths = c(0,diff(`Cumulative Count`, 1))) 



# Import Scottish covid data
#path <- "COVID-19_Scotland_data_all_2020-04-10.xlsx" 
path <- paste0("COVID-19_Scotland_data_all_", {Sys.Date()}, ".xlsx") 
# Scottish cases data
scot_data_raw <- read_excel(path, sheet = "Cases By Health Board", skip = 1)
scot_data_raw <- filter(scot_data_raw, !Health_Board %in% c("Increase", "pIncrease")) %>%
              rename(confirmed_cases= Total,
                     Date = Health_Board) %>%
              mutate(date = lubridate::ymd(Date)) %>%
              select(-Date)
  
scot_data_raw$date[nrow(scot_data_raw)] <- scot_data_raw$date[nrow(scot_data_raw)-1] + days(1)
#scot_data_raw$confirmed_cases[nrow(scot_data_raw)] <- 4565 ## REMOVE

scot_data <- scot_data_raw %>%
             mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases),0)) %>%
             mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) 
             
# Per health board
scot_data_health_board <- scot_data %>% 
  select(date, Ayrshire:`Dumfries and Galloway`) %>%
  pivot_longer(Ayrshire:`Dumfries and Galloway`,
               names_to = "health_board",
               values_to = "confirmed_cases") %>% 
  group_by(health_board) %>%
  mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases), 0)) %>%
  ungroup() %>%
  replace_na(list(new_cases = 0))

# Per health board for map
# Cumulative incidence
scot_ci_hb <- read_excel(path, sheet = "Cumulative Incidence Grouped") %>%
  slice(nrow(.)) %>%
  select(Ayrshire:Tayside) %>%
  pivot_longer(Ayrshire:Tayside, names_to = "health_board", values_to = "cumulative_incidence")

# Incidence over last day
scot_ti_hb <- read_excel(path, sheet = "Incidence by Health Board") %>%
  slice(nrow(.)) %>%
  select(Ayrshire:Tayside) %>%
  pivot_longer(Ayrshire:Tayside, names_to = "health_board", values_to = "today_incidence")

scot_data_health_board_total <- scot_data_health_board %>% 
                                group_by(health_board) %>%
                                summarise(confirmed_cases = max(confirmed_cases, na.rm = T)) %>%
                                left_join(scot_ci_hb)

# Scottish death data
scot_deaths <- read_excel(path, sheet = "Scotland Deaths", skip = 1) %>%
  rename("deaths" = Deaths_Cum, 
         "new_deaths" = Deaths_New) %>%
  mutate(doubling_time_week = 7*log(2)/log(deaths/replace_na(lag(deaths,7),0))) %>%
  mutate(date = lubridate::ymd(Date)) 

scot_deaths$date[nrow(scot_deaths)] <- scot_deaths$date[nrow(scot_deaths)-1] + days(1)

# Scottish tests
scot_tests <- read_excel(path, sheet = "CPT & DPC") 
#scot_tests$Cases[nrow(scot_tests)] <- 4565 #REMOVE

scot_tests  <- scot_tests %>%
  rename("Conducted" = Tests, 
         "Total Positive" = Cases,
         "deaths_per_case" = DPC,
         "cases_per_test" = CPT) %>%
  mutate("Conducted today" = Conducted - replace_na(lag(Conducted), 0)) %>%
  mutate("Total Negative" = Conducted - `Total Positive`) %>%
  mutate("Positive" = `Total Positive` - replace_na(lag(`Total Positive`), 0),
         "Negative" = `Total Negative` - replace_na(lag(`Total Negative`), 0)) 
scot_tests$Positive[scot_tests$Date == ymd("2020-03-12")] <- 24
scot_tests$Negative[scot_tests$Date == ymd("2020-03-12")] <- 552

  
scot_tests_long <- scot_tests %>%
pivot_longer(cols = Positive:Negative, names_to = "Result", values_to = "Number") %>%
mutate(Result = factor(Result, levels = c("Positive", "Negative")))


# Map files
# SCOTLAND MAP
cases_by_area <- sf::st_read("SG_NHS_HealthBoards_2019b.geojson") %>%
                 mutate(health_board = case_when(HBName == "Ayrshire and Arran" ~ "Ayrshire",
                                                 HBName %in% c("Grampian", "Shetland", "Orkney") ~ "Grampian, Shetland and Orkney",
                                                 HBName %in% c("Highland", "Western Isles") ~ "Highland and Western Isles",
                        TRUE ~ as.character(HBName))) %>%
                 st_transform(crs = st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")) %>%
                 left_join(scot_data_health_board_total, by = c("health_board" = "health_board")) 

# Plot colours
#nice_blue <- palette_OkabeIto[5]
#nice_red <- palette_OkabeIto[6]  

nice_blue <- "#0072B2"
nice_red <- "#D55E00"
nice_green <- "#009E73"
nice_orange <- "#E69F00"

# Weekly Growth 
wk_gr_newcases <- scot_data %>% select(new_cases, date) %>%
      mutate(outcome = new_cases) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
       mutate(adjustment = "no adjustment")

wk_gr_newdeaths <- scot_deaths %>% select(new_deaths, date) %>%
      mutate(outcome = new_deaths) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk()

wk_gr_newdeaths_nrs <- nrs_deaths_covid_raw_daily %>% select(new_deaths, date) %>%
      mutate(outcome = new_deaths) %>%
       weekly_ratios() %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk()

wk_gr_newdeaths_nrs_hps <- bind_rows(wk_gr_newdeaths %>% mutate(source = "HPS"),
wk_gr_newdeaths_nrs %>% mutate(source = "NRS"))

wk_gr_newcases_hb <- scot_data_health_board %>% rename("outcome" = new_cases) %>%
   group_by(health_board) %>%
   nest() %>%
   mutate(gr = map(data, ~weekly_ratios(.x))) %>%
   unnest(gr) %>%
   filter(date >= ymd("2020-03-23")) %>%
       filter(!is.na(ratio) & !is.infinite(ratio)) %>%
       significance_wk() %>%
    mutate(symbol = case_when(significance == "not significantly different to 1" ~ "not sig <>1",
                            significance == "not significantly different to 1" ~ "not sig <>1",
                            significance == "significantly greater than 1" ~ "sig <> 1",
                            significance == "significantly less than 1" ~ "sig <> 1")) 

nrs_deaths_covid_week <- nrs_deaths_covid_raw %>% 
  filter(`Week beginning` == "Deaths involving COVID-194" | 
          Details %in% c("Care Home", "Home / Non-institution", "Hospital", "Other institution")) %>%
  pivot_longer(cols = `43829`:names(nrs_deaths_covid_raw)[ncol(nrs_deaths_covid_raw)-1],
               names_to = "week_beginning", 
               values_to = "weekly_deaths_total") %>%
  mutate(Details = case_when(`Week beginning` == "Deaths involving COVID-194" ~ "All",
                             TRUE ~ Details)) %>%
  mutate(week_beginning = janitor::excel_numeric_to_date(parse_number(week_beginning))) %>%
  mutate(Type = "Covid19 deaths") %>%
  rename("Location" = Details) %>%
  filter(Location != "Other institution") %>%
  select(Location, week_beginning, weekly_deaths_total, Type) %>%
  mutate(week_ending = week_beginning + 6)

wk_gr_nrs_deaths <- nrs_deaths_covid_week %>%
  group_by(Location) %>%
  mutate(week_ending_prev = lag(weekly_deaths_total, 1)) %>%
  filter(!is.na(week_ending_prev) & week_ending_prev >0) %>% 
  pivot_longer(c(weekly_deaths_total,week_ending_prev), names_to = "week", values_to = "new_numbers_week") %>%
  mutate(week = factor(week, levels = c("week_ending_prev", "weekly_deaths_total"))) %>%
  ungroup() %>%
  group_by(Location, week_ending) %>%
  nest() %>%
  mutate(mod = map(data, ~ glm(new_numbers_week ~ week,
                              data = .x, family = poisson))) %>%
  mutate(ratio_m = map_dbl(mod, ~ exp(coef(.x)["weekweekly_deaths_total"]))) %>% 
  mutate(lci = map_dbl(mod, ~ exp(confint(.x)["weekweekly_deaths_total", "2.5 %"]))) %>%
  mutate(uci = map_dbl(mod, ~ exp(confint(.x)["weekweekly_deaths_total", "97.5 %"]))) %>%
  ungroup() %>%
  select(Location, week_ending, ratio_m, lci, uci) %>%
  mutate(comparison = if_else(ratio_m >1, "greater than", "less than")) %>%
  significance_wk() %>%
  rename(date = week_ending)

wk_gr_nrs_deaths <- left_join(wk_gr_nrs_deaths,
  nrs_deaths_covid_week %>% 
  filter(weekly_deaths_total >0) %>%
  rename("change" = weekly_deaths_total, 
         "date" = "week_ending") %>%
  select(change, date, Location))

wk_gr_newdeaths_latest <- wk_gr_newdeaths %>% mutate(source = "HPS", Location = "HPS") %>% filter(date == max(wk_gr_newdeaths$date))

wk_gr_nps_nrs <- bind_rows(wk_gr_newdeaths %>% mutate(source = "HPS")  %>%
                           filter(date %in% wk_gr_nrs_deaths$date), 
                           wk_gr_nrs_deaths %>% mutate(source = "NRS"))
wk_gr_nps_nrs <- wk_gr_nps_nrs %>%
                 mutate(Location = case_when( source == "HPS" ~ source, 
                                              Location == "All" ~ "NRS",
                                             TRUE ~ Location))
  
```


Numbers of new cases and new deaths, as reported by Health Protection Scotland (HPS), have been monitored on a daily basis (see Figure \@ref(fig:traj)). To assess the rate of spread of COVID-19 in Scotland, we have considered the ratio of the sum of these metrics in the past week compared to their respective sums in the previous week. A similar ratio has been used to measure changes in hospital bed occupancy by comparing daily numbers of bed and ICU occupancy to those one week ago (see Figure \@ref(fig:deathscaseshospicu)).



\subsection*{Data}
The cumulative number of cases and deaths has been reported each day since the start of the pandemic in Scotland by Health Protection Scotland (HPS). The numbers include deaths which have been registered with National Records of Scotland (NRS) where a laboratory confirmed report of COVID-19 in the 28 days prior to death exists. These numbers are expected to capture the majority of deaths occuring in hospitals, but a lower proportion of those in care homes and the community. 

The National Records of Scotland (NRS) also provide a weekly report of all deaths where COVID-19 is mentioned on the death certificate (not just those confirmed by a test as provided by HPS). The report provides information by place of death, location, age and gender. 

Bed occupancy numbers in Scottish hospitals are reported each day by NHS Boards. The number of hospital beds occupied by COVID-19 patients and the number of these patients in intensive care units (ICU) are provided.

Several caveats are listed below which should be considered when interpreting trends in the reported numbers.

\subsection*{Weekly comparisons}
We have used ratios of both cases and deaths in the past week compared to the previous week (i.e. the total over the 7 days as a ratio of the total over the previous 7 days) to monitor the weekly rate of spread of COVID-19 (see Figures \@ref(fig:ratiosnrshpsloc), \@ref(fig:ratiosnrshpsdaily) and \@ref(fig:deathscaseshospicu)). Weekly totals are considered as they are less prone to week day variations than the daily values (see Caveats below).

Ratios for hospital occupancy are calculated as the number of patients in hospital (or ICU) on a given day compared to the number of patients in hospital (or ICU) one week before (e.g. the ratio for 28th April is ‘occupancy on 28th April’ divided by ‘occupancy on 21st April’).

The confidence interval for the log of the ratio of two frequencies (w1/w2) may be approximated by {1/w1 + 1/w2}, based on assuming a Poisson distribution for w1 and w2 (w1 = total cases (or deaths) in past week and w2 = total cases (or death) in previous week, or for bed occupancy w1=beds occupied on the given day, w2=beds occupied on that day the week before). The approximation is obtained from the standard formula: 
\hfill \break


\begin{tabular}{rrll}\
& $Var\{log(a)\}$ & $\approx$ & $Var(a)/Mean(a)^2$\\
&  & = & $a/a^2$ for a Poisson frequency \\
& & = & $1/a$ \\
& & & \\
\multicolumn{2}{c}{The ratio on a log scale may be expressed:} & &\\
& & & \\
& $log(w1/w2)$ & = & $log(w1)$ – $log(w2)$ \\
& & & \\
 \multicolumn{2}{r}{giving} & &\\
& & & \\
& $Var\{log(w1/w2)\} $ & = & $Var\{log(w1)$ - $log(w2)\}$\\
& & = & $Var\{log(w1)\}$ + $Var\{log(w2)\}$  \\
&  & = &  $\{1/w1 + 1/w2\}$ \\
& & & \\
 \multicolumn{2}{r}{and} & &\\
&  & & \\
& $SE\{log(w1/w2)\}$ & = & $\sqrt\{1/w1$ + $1/w2\}$ \\
\end{tabular}


\hfill \break
The 95% confidence interval is then calculated as log(w1/w2) ± 1.96 x SE, which is exponentiated to provide the confidence intervals displayed.

\subsection*{Caveats}

We list several caveats which should be borne in mind when considering the statistics presented:

Daily numbers of cases and deaths are affected by:

-	Changes in testing policy which in turn is affected by the availability of tests and laboratory facilities for their analysis. For example, testing of people in care homes and healthcare workers is being increased. 

-	Hospital admission policy – if admission criteria are relaxed then a higher proportion of Covid-19 patients are admitted and tested and hence included in the daily case and (HPS) death statistics. 

-	Differences in reporting of deaths to NRS and HPS due to variable factors such as lag caused by weekends and public holidays. 

HPS death counts include individuals who are tested outside hospital. An increased frequency of testing in care homes coupled with the increasing number of deaths in this setting (see Figures \@ref(fig:deathsno) and \@ref(fig:ratiosnrshpsloc)) makes HPS numbers difficult to compare over time. The NRS ‘all COVID-19’ death numbers may allow a more consistent comparison to be made, and additionally allow care home and other non-hospital deaths to be considered separately. However, these data are only issued on a weekly basis and they will be less up-to-date than that from HPS.

Given the later availability of NRS data and the variable factors affecting both the number of cases and HPS deaths, some consideration might be given to hospital occupancy data (see Figure \@ref(fig:deathscaseshospicu)). However, these data will not reflect cases in care homes, and a consistent criteria for admission to hospital or to ICU would need to be assumed in order to make comparisons over time.

```{r}
nrs_hosp <- filter(wk_gr_nps_nrs, Location =="Hospital" & date == max(wk_gr_nps_nrs$date))
nrs_carehomes <- filter(wk_gr_nps_nrs, Location =="Care Home" & date == max(wk_gr_nps_nrs$date))
```


The ratios by place of death (Figure \@ref(fig:ratiosnrshpsloc)) will be subject to variations in hospital admission policy. For example, if a greater proportion of care home residents are being admitted to hospital then, under an unchanged admission policy, we might expect the weekly ratio for care homes to be higher than the last reported value of `r round(nrs_carehomes$ratio_m,2)` (95% CI: `r round(nrs_carehomes$lci,2)`-`r round(nrs_carehomes$uci,2)`), and the weekly ratio for hospitals to be lower than the reported value of `r round(nrs_hosp$ratio_m,2)` (95% CI: `r round(nrs_hosp$lci,2)`-`r round(nrs_hosp$uci,2)`) (data for week ending `r as.character(max(wk_gr_nps_nrs$date))`).

The confidence interval for the ratios indicates the accuracy of the observed ratio of the two particular weeks considered. However, there are many factors which influence the ratios other than the true underlying rate of spread of COVID-19. Some of these will be systematic (as described in the above caveats) and others may be more random. The caveats should all be considered when interpreting the ratios, particularly when the confidence interval does not include 1. A modelling approach would be suggested to confirm whether the ratios have consistently fallen below 1 and to determine whether their values have stabilised. Note that while the ratios are related to the R number and ratios less than (or greater than) 1 are expected to correspond to R numbers of less than (or greater than) 1, their values will not be the same. 

\hfill \break
\hfill \break

```{r, fig.width=6, fig.cap="Ratio of sum of new cases during current week compared to previous week. Comparison between latest estimate of NRS and HPS data", eval=FALSE}
wk_gr_nps_nrs %>% 
filter(Location %in% c("NRS", "HPS")) %>%
filter(date == max(wk_gr_nps_nrs$date)) %>% 
ggplot(aes(x = Location, y = ratio_m,
           ymin = lci, ymax = uci, 
    colour = Location, fill = Location)) + 
  geom_point(size = 2) +
  geom_errorbar(width = 0, alpha = 0.6, size = 1.5) + 
  labs( x = "Data source",
        y = "Weekly ratio",
        title = "Change in weekly deaths total as ratio of previous week", 
        subtitle = str_glue("Date: ", {as.character(max(wk_gr_nps_nrs$date))} ), 
        colour = "Location/Source",
        fill = "Location/Source") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```


```{r traj, fig.width=6, fig.height=5, fig.cap="Trajectories of COVID-19 cases and deaths vs respective cumulative numbers. Data source: HPS" }
cases_deaths <- bind_rows(select(scot_data, date, new_cases, confirmed_cases) %>% 
            rename("new" = new_cases, "cumulative" = confirmed_cases) %>%
              mutate(type = "cases") %>%
              mutate(change = roll_sumr(new, n = 7)) ,
         select(scot_deaths, date, new_deaths, deaths) %>% 
           filter(deaths >0) %>%
            rename("new" = new_deaths, "cumulative" = deaths) %>%
              mutate(type = "deaths") %>%
              mutate(change = roll_sumr(new, n = 7)))

arrow_date <- max(cases_deaths$date)-15
#arrow_date <- ymd("2020-03-23")
cases_deaths_date <- subset(cases_deaths, date == arrow_date) %>%
                     mutate(y = change + c(50, 20),
                            yend = change + c(500, 100), 
                            yend_date = yend + c(122, 40))


ggplot(cases_deaths) +
    aes(x = cumulative, y = change, group = type, colour = type) +
    geom_line(lwd = 1) +
    geom_point(size = 0.7) +
    scale_x_continuous(trans = "log10",
                       labels = scales::comma_format(accuracy = 1),
                       #breaks = breaks,
                       expand = expansion(mult = 0.01), 
                       limits = c(10, NA)) +
    scale_y_continuous(trans = "log10",
                       labels = scales::comma_format(accuracy = 1),
                       limits = c(min(filter(cases_deaths, cumulative >=10)$change, na.rm = TRUE), max(cases_deaths$change, na.rm = TRUE)+500),
                       #expand = expansion(mult = 0.05)
                       ) +
    geom_segment(data = cases_deaths_date, colour = "black", aes(x = cumulative, y = y, xend = cumulative, yend = yend),
               arrow = arrow(length = unit(2, "mm"), ends = "first")) +
    geom_text(data = cases_deaths_date, colour = "black", aes(x = cumulative, y = yend_date, label = date)) +
    guides(x = guide_axis(check.overlap = TRUE), y = guide_axis(check.overlap = TRUE)) +
    theme(legend.position = "none") +
    labs(y = str_glue("New cases/deaths \n(in the past 7 days)"),
         x = str_glue("Total Number"),
         title = str_glue("Trajectory of COVID-19 cases/deaths vs current total"), 
         #subtitle = str_glue("Arrows = ", {as.character(arrow_date)}),
         subtitle = str_glue("Starting date: ", {as.character(min(filter(cases_deaths, cumulative >=10)$date))}),
         colour = "") +
  theme_bw(base_size = 12) +
  theme(legend.position = c(0.20, 0.85))
```



\newpage 

```{r deathsno, fig.width=8,fig.cap="Number of deaths per week, compared by location and data source "}
ggplot(bind_rows(wk_gr_nps_nrs, wk_gr_newdeaths_latest) %>%
         mutate(Location = factor(Location, levels = c("HPS", "NRS","Care Home", "Hospital","Home / Non-institution")))) +
  aes(x = date, y = change,
    colour = Location, fill = Location) +
  geom_line(position=position_dodge(width=1)) +
  geom_point(size = 1, position=position_dodge(width=1)) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "Week Ending",
        y = "Weekly Deaths",
        title = "Weekly deaths by location and data source", 
        colour = "Location/Source",
        fill = "Location/Source") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw(base_size = 12)
```

```{r deathsnodaily, fig.width=8,fig.cap="Cumulative number of deaths, compared by data source "}
ggplot(bind_rows(scot_deaths %>% mutate(Source = "HPS"), nrs_deaths_covid_raw_daily %>% rename(deaths = `Cumulative Count`)) ) +
  aes(x = date, y = deaths,
    colour = Source, fill = Source) +
  geom_line() +
  geom_point(size = 1) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "Date",
        y = "Cumulative Deaths",
        title = "Cumulative number of deaths by data source", 
        colour = "Source",
        fill = "Source") +
  scale_x_date(date_breaks = "2 week", labels = date_format("%b %d")) +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw(base_size = 12)
```




```{r ratiosnrshpsloc, fig.width=8,fig.cap="Ratio of new deaths during current week vs previous week, compared by location and data source"}

fig1 <- 
  bind_rows(wk_gr_nps_nrs, wk_gr_newdeaths_latest) %>%
  mutate(Location = factor(Location, levels = c("HPS", "NRS","Care Home", "Hospital", "Home / Non-institution"))) %>%
  ggplot() +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci, 
    colour = Location, fill = Location) +
  geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=1)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(position=position_dodge(width=1)) +
  geom_point(size = 1, position=position_dodge(width=1)) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        title = "Weekly ratio of new deaths", 
        colour = "Location/Source",
        fill = "Location/Source") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw(base_size = 12)


fig1_inset <- wk_gr_nps_nrs %>% 
filter(Location %in% c("NRS", "HPS")) %>%
filter(date == max(wk_gr_nps_nrs$date)) %>% 
ggplot(aes(x = Location, y = ratio_m,
           ymin = lci, ymax = uci, 
    colour = Location, fill = Location)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_point(size = 1.8) +
  geom_errorbar(width = 0, alpha = 0.6, size = 1.2) + 
  labs( x = "",
        y = "",
        #title = "Change in weekly deaths total as ratio of previous week", 
        subtitle = str_glue("Date: ", {as.character(max(wk_gr_nps_nrs$date))} ), 
        colour = "Location/Source",
        fill = "Location/Source") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw(base_size = 10) +
  theme(legend.position = "none") +
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))

ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.45, y = .48, width = .23, height = .35)
```



```{r ratiosnrshps, fig.cap="Ratio of new deaths during current week vs previous week, compared by data source", eval = FALSE}
ggplot(wk_gr_nps_nrs %>% filter(Location %in% c("NRS", "HPS"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci, 
    colour = Location, fill = Location) +
  geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=0.3)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.3) +
  geom_line(position=position_dodge(width=0.3)) +
  geom_point(size = 1, position=position_dodge(width=0.3)) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        title = "Change in weekly deaths total as ratio of previous week",
        colour = "Source",
        fill = "Source") +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw(base_size = 12)
```

```{r ratiosnrshpsdaily, fig.width=8, fig.cap="Ratio of new deaths during current week vs previous week, compared by data source"}
fig1<- ggplot(wk_gr_newdeaths_nrs_hps %>% filter(!is.na(ratio) & date >= dmy("23/03/2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci, 
    colour = source, fill = source) +
  geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=0.3)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.3) +
  geom_line(position=position_dodge(width=0.3)) +
  geom_point(size = 1, position=position_dodge(width=0.3)) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly ratio",
        title = "Weekly ratio of new deaths",
        colour = "Source",
        fill = "Source") +
  scale_x_date(date_breaks = "2 week", labels = date_format("%b %d")) +
  scale_colour_brewer(palette = "Dark2") +
  theme_bw(base_size = 12)

fig1_inset <- ggplot(wk_gr_newdeaths_nrs_hps %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths_nrs_hps$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci,
    colour = source, fill = source) +
  geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=0.2)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(position=position_dodge(width=0.2)) +
  geom_point(size = 1, position=position_dodge(width=0.2)) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  scale_colour_brewer(palette = "Dark2") +
  labs( x = "", y = "") +
  theme_bw(base_size = 10) +
  theme(legend.position = "none") 

ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.40, y = .50, width = .42, height = .36) 
```

```{r nrshpsdailydeathsratios, fig.width=6, fig.height= 6, fig.cap="Ratio of new deaths vs respective counts from previous week, compared by data source", eval = FALSE}

fig1 <- ggplot(wk_gr_newdeaths_nrs %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New Deaths (Source: NRS)") +
  theme_bw(base_size = 12) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 11, b = 0, l = 0)))

fig1_inset <- ggplot(wk_gr_newdeaths_nrs %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths_nrs$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 10)

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.52, y = .48, width = .46, height = .4)

fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 12)

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10)

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.52, y = .48, width = .46, height = .40)


gridExtra::grid.arrange(fig1, fig2, nrow = 2) 
```


\newpage

```{r casesdeathsratios, fig.width=6, fig.height= 6, fig.cap="Ratio of new cases and deaths vs respective counts from previous week", eval = FALSE}
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New cases") +
  theme_bw(base_size = 12) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 11, b = 0, l = 0)))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 10)

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.52, y = .48, width = .46, height = .4)

fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "End Date of Current Weekk",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 12)

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10)

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.52, y = .48, width = .46, height = .40)


gridExtra::grid.arrange(fig1, fig2, nrow = 2) 
```


\newpage

```{r icu, fig.width=6, fig.height= 6, fig.cap="Week over week comparison of ICU occupancy and hospitalisations, measured on a daily basis", eval=FALSE}

fig1 <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Hospitalisations (Source: NHS Boards)") +
  theme_bw(base_size = 12)


fig1_inset <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_hosp$date)-8)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.52, y = .48, width = .46, height = .40)


fig2 <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "ICU patients") +
  theme_bw(base_size = 12)



fig2_inset <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_icu$date)-8)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10)

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.52, y = .48, width = .46, height = .40)

gridExtra::grid.arrange(fig1, fig2, nrow = 2) 
```

\newpage
\thispagestyle{empty}

```{r deathscaseshospicu, fig.width=5.5, fig.height= 9, fig.cap="Weekly ratios for new cases and new deaths are ratios of weekly sums, plotted on a log10 scale. Day-of-the-week ratios for hospital occupancy and ICU occupancy are ratios of daily counts, measured 7 days apart and plotted on a log10 scale. Figure insets highlight more recent ratios and trends."}
fig1 <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        #title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New Cases (Source: HPS)") +
  theme_bw(base_size = 12) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 14, b = 0, l = 0)))

fig1_inset <- ggplot(wk_gr_newcases %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newcases$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_blue, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_blue) +
  geom_point(colour = nice_blue, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "", y = "") +
  theme_bw(base_size = 10) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig1 <- ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.57, y = .52, width = .40, height = .3)

fig2 <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
              !is.infinite(ratio) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Deaths (Source: HPS)") +
  theme_bw(base_size = 12)

fig2_inset <- ggplot(wk_gr_newdeaths %>% filter(!is.na(ratio) & 
                    !is.infinite(ratio) & date >= max(wk_gr_newdeaths$date)-8)) +
  aes(x = date, y = ratio,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_red, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_red) +
  geom_point(colour = nice_red, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"),expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig2 <- ggdraw() +
  draw_plot(fig2) +
  draw_plot(fig2_inset, x = 0.57, y = .52, width = .40, height = .3)


fig3 <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "ICU Occupancy (Source: NHS Boards)") +
  theme_bw(base_size = 12) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 7, b = 0, l = 0)))



fig3_inset <- ggplot(wk_gr_hosp_icu %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_icu$date)-8)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_orange, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_orange) +
  geom_point(colour = nice_orange, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig3 <- ggdraw() +
  draw_plot(fig3) +
  draw_plot(fig3_inset, x = 0.57, y = .52, width = .40, height = .3)

fig4 <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
              !is.infinite(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_y_continuous(trans = "log10",
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  labs( x = "Current Date",
        y = "Day-Of-The-Week Ratio",
        subtitle = "Hospital Occupancy (Source: NHS Boards)") +
  theme_bw(base_size = 12) +
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 14, b = 0, l = 0)))


fig4_inset <- ggplot(wk_gr_hosp_hosp %>% filter(!is.na(ratio_m) & 
                    !is.infinite(ratio_m) & date >= max(wk_gr_hosp_hosp$date)-8)) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci) +
  geom_errorbar(width = 0, colour = nice_green, alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line(colour = nice_green) +
  geom_point(colour = nice_green, size = 1) +
  scale_x_date(date_breaks = "4 days", labels = date_format("%b %d"), expand = expansion(mult = 0.1)) +
  scale_y_continuous(trans = "log10") +
  labs( x = "",
        y = "") +
  theme_bw(base_size = 10) +
  theme(plot.margin = margin(0, 0, -1, 0, "cm"))

fig4 <- ggdraw() +
  draw_plot(fig4) +
  draw_plot(fig4_inset, x = 0.57, y = 0.52, width = 0.40, height = 0.3)

gridExtra::grid.arrange(fig1, fig2, fig3, fig4, nrow = 4) 
```


\newpage

```{r, fig.width=6, eval = FALSE}
p<-ggplot(wk_gr_newcases_hb) +
  aes(
    x = date, y = ratio,
    #ymin = lci, ymax = uci,
    colour = health_board, fill = health_board
  ) +
  #geom_errorbar(width = 0, colour = "gray30", alpha = 0.4) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line() +
  geom_point(aes(shape = symbol), size = 2) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  scale_color_manual(values = c('#e6194b', '#ffe119', '#4363d8','#800000', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#008080', '#9a6324',  '#808000', '#ffd8b1', '#000075', '#808080'))+
  scale_fill_manual(values = c('#e6194b', '#ffe119', '#4363d8','#800000', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#008080', '#9a6324',  '#808000', '#ffd8b1', '#000075', '#808080'))+
  scale_shape_manual(values = c(1,16)) +
  labs( x = "End Date of Current Week",
        y = "Weekly Ratio",
        subtitle = "New Cases") +
  guides(colour = guide_legend(override.aes = list(shape = NA), nrow = 6),
         shape = guide_legend(nrow = 6)
         ) +
  theme_bw(base_size = 12) +
    theme(legend.position="top",
        legend.box="horizontal", legend.margin=margin(),
        legend.title = element_blank(),
        legend.text=element_text(size=8)) 

p


```

```{r deathdayweek, fig.width=7, fig.cap="Number of deaths by weekday, shown for different weeks"}
first_day <- as.data.frame(cbind(first_day = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
levels = list(list("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
list("Tue", "Wed", "Thu", "Fri", "Sat", "Sun", "Mon"),
list("Wed", "Thu", "Fri", "Sat", "Sun","Mon", "Tue"),
list("Thu", "Fri", "Sat", "Sun", "Mon", "Tue", "Wed"),
list("Fri", "Sat", "Sun", "Mon", "Tue", "Wed", "Thu"),
list("Sat", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri"),
list("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))))


scot_deaths_week <- scot_deaths %>%
  filter(date > ymd("2020-03-12")) %>%
mutate(dayofweek = weekdays(date, abbreviate = TRUE)) %>%
mutate(day_no = case_when(dayofweek == "Mon" ~ 0, 
                          dayofweek == "Tue" ~ 6,
                          dayofweek == "Wed" ~ 5,
                          dayofweek == "Thu" ~ 4,
                          dayofweek == "Fri" ~ 3,
                          dayofweek == "Sat" ~ 2,
                          dayofweek == "Sun" ~ 1
                          )) %>%
#mutate(week = cut(date + 1, "week")) %>%
mutate(week = cut(date + day_no[nrow(.)-6], "week")) %>%
mutate(week = ymd(week)-day_no[nrow(.)-6])%>%
mutate(dayofweek = factor(dayofweek, 
                          levels = unlist(subset(first_day, first_day == .$dayofweek[nrow(.)-6])$levels)))


scot_data_week <- scot_data %>%
mutate(dayofweek = weekdays(date, abbreviate = TRUE)) %>%
mutate(day_no = case_when(dayofweek == "Mon" ~ 0, 
                          dayofweek == "Tue" ~ 6,
                          dayofweek == "Wed" ~ 5,
                          dayofweek == "Thu" ~ 4,
                          dayofweek == "Fri" ~ 3,
                          dayofweek == "Sat" ~ 2,
                          dayofweek == "Sun" ~ 1
                          )) %>%
#mutate(week = cut(date + 1, "week")) %>%
mutate(week = cut(date + day_no[nrow(.)-6], "week")) %>%
mutate(week = ymd(week)-day_no[nrow(.)-6])%>%
mutate(dayofweek = factor(dayofweek, 
                          levels = unlist(subset(first_day, first_day == .$dayofweek[nrow(.)-6])$levels)))


last_two_weeks_deaths <- unique(arrange(scot_deaths_week, desc(date))$week)[1:2] 
last_two_weeks <- unique(arrange(scot_data_week, desc(date))$week)[1:2] 



ggplot(scot_deaths_week) + 
  geom_point(aes(x = dayofweek, y = new_deaths, colour = factor(week))) +
  geom_line(aes(x = dayofweek, y = new_deaths, colour = factor(week), group = week)) + 
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) + theme_bw(base_size = 12) +
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "", y = "New deaths", colour = "Week beginning")
```


```{r, fig.width=7, fig.cap="Number of cases by weekday, shown for different weeks"}
ggplot(scot_data_week) + 
  geom_point(aes(x = dayofweek, y = new_cases, colour = factor(week))) +
  geom_line(aes(x = dayofweek, y = new_cases, colour = factor(week), group = week)) + 
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) + theme_bw(base_size = 12) +
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "", y = "New cases", colour = "Week beginning")
```

```{r, fig.width=6, fig.height=3.5, fig.cap="Number of cases and deaths by week day, shown for the last two weeks. Note that the distribution of cases and deaths in w/b 2020-04-13 may be affected by an extended reporting lag due to Easter weekend."}
# ggplot(scot_data_week %>% filter(week %in% last_two_weeks)) + 
#   geom_point(aes(x = dayofweek, y = new_cases, colour = factor(week))) +
#   geom_line(aes(x = dayofweek, y = new_cases, colour = factor(week), group = week)) + 
#   scale_y_continuous(trans = "log10",
#                          labels = scales::comma_format(accuracy = 1),
#                          #breaks = breaks,
#                          expand = expansion(mult = 0.1)) + theme_bw(base_size = 12) +
#   scale_colour_brewer(palette = "Dark2") +
#   guides(colour = guide_legend(reverse = TRUE)) +
#   labs(x = "", y = "New cases", colour = "Week beginning")

bind_rows(scot_data_week %>% filter(week %in% last_two_weeks) %>% mutate(Type = "Cases") %>% rename("New" = new_cases),
          scot_deaths_week %>% filter(week %in% last_two_weeks) %>% mutate(Type = "Deaths") %>% rename("New" = new_deaths)) %>% 
  ggplot() + 
  geom_point(aes(x = dayofweek, y = New, colour = factor(week))) +
  geom_line(aes(x = dayofweek, y = New, colour = factor(week), group = week)) + 
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) + theme_bw(base_size = 12) +
  scale_colour_brewer(palette = "Dark2") +
  guides(colour = guide_legend(reverse = TRUE)) +
  labs(x = "", y = "Number of new cases or deaths", colour = "Week beginning") + facet_grid(Type~., scales = "free_y")
  
```



```{r, eval = FALSE}
all <- bind_rows(wk_gr_newcases, wk_gr_newcases_byday)
fig1 <- ggplot(all %>% filter(!is.na(ratio_m) & date >= dmy("23-03-2020"))) +
  aes(x = date, y = ratio_m,
    ymin = lci, ymax = uci, 
    colour = adjustment, fill = adjustment) +
  geom_errorbar(width = 0, alpha = 0.4, position=position_dodge(width=0.5)) +
  geom_hline(yintercept = 1, colour = "red", linetype = "dashed", size = 0.2) +
  geom_line() +
  geom_point(size = 1, position=position_dodge(width=0.3)) +
  scale_y_continuous(trans = "log10",
                         labels = scales::comma_format(accuracy = 1),
                         #breaks = breaks,
                         expand = expansion(mult = 0.1)) +
  labs( x = "End date of current week",
        y = "Weekly ratio",
        title = "Change in weekly cases/deaths total as ratio of previous week",
        subtitle = "New cases") +
  theme_bw(base_size = 12)
fig1
```


This report was produced by members of Epigroup and The Roslin Institute at the University of Edinburgh.


