---
output: word_document
urlcolor: blue
---

```{r setup, include=FALSE}
#date: "`r format(Sys.time(), '%d %B, %Y')`" 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Load packages
library(flexdashboard) ; library(shiny) ; library(readr); library(dplyr); library(tidyr); library(purrr); library(forcats); library(stringr); library(htmlwidgets); library(lubridate); library(sf); library(RcppRoll); library(plotly); library(shinythemes);library(leaflet); library(classInt); library(ggrepel); library(scales); library(leaflet.extras); library(RColorBrewer);
library(colorblindr); library(readxl)


# Import Scottish covid data
#path <- "COVID-19_Scotland_data_all_2020-04-10.xlsx" 
path <- paste0("COVID-19_Scotland_data_all_", {Sys.Date()}, ".xlsx")
# Scottish cases data
scot_data_raw <- read_excel(path, sheet = "Cases By Health Board", skip = 1)
scot_data_raw <- filter(scot_data_raw, !Health_Board %in% c("Increase", "pIncrease")) %>%
              rename(confirmed_cases= Total,
                     Date = Health_Board) %>%
              mutate(date = lubridate::ymd(Date)) %>%
              select(-Date)
  
scot_data_raw$date[nrow(scot_data_raw)] <- scot_data_raw$date[nrow(scot_data_raw)-1] + days(1)
#scot_data_raw$confirmed_cases[nrow(scot_data_raw)] <- 4565 ## REMOVE

scot_data <- scot_data_raw %>%
             mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases),0)) %>%
             mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) 
             
# Per health board
scot_data_health_board <- scot_data %>% 
  select(date, Ayrshire:`Dumfries and Galloway`) %>%
  pivot_longer(Ayrshire:`Dumfries and Galloway`,
               names_to = "health_board",
               values_to = "confirmed_cases") %>% 
  group_by(health_board) %>%
  mutate(new_cases = confirmed_cases - replace_na(lag(confirmed_cases), 0)) %>%
  ungroup() %>%
  replace_na(list(new_cases = 0))

# Per health board for map
# Cumulative incidence
scot_ci_hb <- read_excel(path, sheet = "Cumulative Incidence Grouped") %>%
  slice(nrow(.)) %>%
  select(Ayrshire:Tayside) %>%
  pivot_longer(Ayrshire:Tayside, names_to = "health_board", values_to = "cumulative_incidence")

# Incidence over last day
scot_ti_hb <- read_excel(path, sheet = "Incidence by Health Board") %>%
  slice(nrow(.)) %>%
  select(Ayrshire:Tayside) %>%
  pivot_longer(Ayrshire:Tayside, names_to = "health_board", values_to = "today_incidence")

scot_data_health_board_total <- scot_data_health_board %>% 
                                group_by(health_board) %>%
                                summarise(confirmed_cases = max(confirmed_cases, na.rm = T)) %>%
                                left_join(scot_ci_hb)

# Scottish death data
scot_deaths <- read_excel(path, sheet = "Scotland Deaths", skip = 1) %>%
  rename("deaths" = Deaths_Cum, 
         "new_deaths" = Deaths_New) %>%
  mutate(doubling_time_week = 7*log(2)/log(deaths/replace_na(lag(deaths,7),0))) %>%
  mutate(date = lubridate::ymd(Date)) 

scot_deaths$date[nrow(scot_deaths)] <- scot_deaths$date[nrow(scot_deaths)-1] + days(1)

# Scottish tests
scot_tests <- read_excel(path, sheet = "CPT & DPC") 
#scot_tests$Cases[nrow(scot_tests)] <- 4565 #REMOVE

scot_tests  <- scot_tests %>%
  rename("Conducted" = Tests, 
         "Total Positive" = Cases,
         "deaths_per_case" = DPC,
         "cases_per_test" = CPT) %>%
  mutate("Conducted today" = Conducted - replace_na(lag(Conducted), 0)) %>%
  mutate("Total Negative" = Conducted - `Total Positive`) %>%
  mutate("Positive" = `Total Positive` - replace_na(lag(`Total Positive`), 0),
         "Negative" = `Total Negative` - replace_na(lag(`Total Negative`), 0)) 
scot_tests$Positive[scot_tests$Date == ymd("2020-03-12")] <- 24
scot_tests$Negative[scot_tests$Date == ymd("2020-03-12")] <- 552

  
scot_tests_long <- scot_tests %>%
pivot_longer(cols = Positive:Negative, names_to = "Result", values_to = "Number") %>%
mutate(Result = factor(Result, levels = c("Positive", "Negative")))


# Map files
# SCOTLAND MAP
cases_by_area <- sf::st_read("SG_NHS_HealthBoards_2019b.geojson") %>%
                 mutate(health_board = case_when(HBName == "Ayrshire and Arran" ~ "Ayrshire",
                                                 HBName %in% c("Grampian", "Shetland", "Orkney") ~ "Grampian, Shetland and Orkney",
                                                 HBName %in% c("Highland", "Western Isles") ~ "Highland and Western Isles",
                        TRUE ~ as.character(HBName))) %>%
                 st_transform(crs = st_crs("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")) %>%
                 left_join(scot_data_health_board_total, by = c("health_board" = "health_board")) 

# Plot colours
nice_blue <- palette_OkabeIto[5]
nice_red <- palette_OkabeIto[6]  
```



Here is the daily report for COVID-19 in Scotland for `r weekdays(max(scot_data$date))`, `r format(max(scot_data$date), '%d %B %Y')`. For the full dashboard, including comparisons with other countries, please visit [here](https://smazeri.shinyapps.io/Covid19_Scotland "COVID-19 Scotland").


**Deaths:**   
The death toll is currently `r max(scot_deaths$deaths, na.rm = TRUE)`.   
The number of new deaths reported today was `r scot_deaths$new_deaths[nrow(scot_deaths)]`, which represents a `r formatC(round(100*(scot_deaths$deaths[nrow(scot_deaths)]-scot_deaths$deaths[nrow(scot_deaths)-1])/scot_deaths$deaths[nrow(scot_deaths)-1],1), format='f', digits=1 )`% increase on yesterday's total.   
The doubling time for deaths over the past 7 days was `r formatC(round(scot_deaths$doubling_time_week[nrow(scot_deaths)],1), format='f', digits=1 )` days.

**Cases:**   
The total number of cases is currently `r comma(scot_data$confirmed_cases[nrow(scot_data)])`.    
The number of new cases today is `r scot_data$new_cases[nrow(scot_data)]`, which represents a `r formatC(round(100*(scot_data$confirmed_cases[nrow(scot_data)]-scot_data$confirmed_cases[nrow(scot_data)-1])/scot_data$confirmed_cases[nrow(scot_data)-1],1), format='f', digits=1 )`% increase on yesterday's total.   
The doubling time for the number of cases over the past 7 days was `r formatC(round(scot_data$doubling_time_week[nrow(scot_data)],1), format='f', digits=1 )` days.

```{r}
data_incidence <- scot_data_health_board %>%
  group_by(health_board) %>%
  arrange(date) %>%
  mutate(relative_increase_cases = 100*(confirmed_cases - replace_na(lag(confirmed_cases),0))/replace_na(lag(confirmed_cases),0)) %>%
  mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) %>%
  ungroup()


data_incidence_today <- subset(data_incidence, date == max(scot_data_health_board$date))
```


**Cases per capita (measured per 10,000 head of population):**   
The highest incidence over the last day was `r formatC(round(max(scot_ti_hb$today_incidence),1), format='f', digits=1 )`, which occurred in `r subset(scot_ti_hb, today_incidence == max(scot_ti_hb$today_incidence))$health_board`.   
The highest cumulative incidence is `r formatC(round(max(scot_ci_hb$cumulative_incidence),1), format='f', digits=1 )`, which is in `r subset(scot_ci_hb, cumulative_incidence == max(scot_ci_hb$cumulative_incidence))$health_board`.

**Health Boards\*:**   
The largest increase in cases in absolute terms by health board was `r max(data_incidence_today$new_cases)` cases, which occurred in `r subset(data_incidence_today, new_cases == max(data_incidence_today$new_cases))$health_board`.   
The largest relative increase in cases was `r formatC(round(max(data_incidence_today$relative_increase_cases),1),format='f', digits=1)`%, which occurred in `r subset(data_incidence_today, relative_increase_cases == max(data_incidence_today$relative_increase_cases))$health_board`.   
The fastest doubling time over the past 7 days was `r formatC(round(min(subset(data_incidence_today, doubling_time_week>0)$doubling_time_week),1),format='f', digits=1)` days, which occurred in `r subset(data_incidence_today, doubling_time_week == min(subset(data_incidence_today, doubling_time_week>0)$doubling_time_week))$health_board`.
<!-- format(round(min(subset(data_incidence_today, doubling_time_week>0)$doubling_time_week)),nsmall = 1)-->

**Tests:**   
The total number of tests reported is `r comma(scot_tests$Conducted[nrow(scot_tests)])`.   
The total number of positive cases to date is `r comma(scot_tests[nrow(scot_tests),"Total Positive"][[1]])`.   
The percentage of tests which were positive is currently `r formatC(round(100*scot_tests[nrow(scot_tests),"Total Positive"]/scot_tests$Conducted[nrow(scot_tests)],1)[[1]],format='f', digits=1 )`%.





\*To enhance graph readability and to account for cases being reallocated across health boards, the following health boards are grouped together: Grampian, Shetland and Orkney; Highland and Western Isles. 

The data in this report are provided by Health Protection Scotland and published daily by the Scottish Government [here](https://www.gov.scot/coronavirus-covid-19/ "gov.scot/coronavirus-covid-19/"). This report has been compiled in collaboration between members of the [Epigroup](https://www.wiki.ed.ac.uk/display/Epigroup/People "Epigroup website") and [EERA](https://www.ed.ac.uk/roslin/eeragroup "EERA website") groups at the University of Edinburgh. 

