---
output: word_document
urlcolor: blue
---

```{r setup, include=FALSE}
#date: "`r format(Sys.time(), '%d %B, %Y')`" 
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
# Load packages
library(flexdashboard) ; library(shiny) ; library(readr); library(dplyr); library(tidyr); library(purrr); library(forcats); library(stringr); library(htmlwidgets); library(lubridate); library(sf); library(RcppRoll); library(plotly); library(shinythemes);library(leaflet); library(classInt); library(ggrepel); library(scales); library(leaflet.extras); library(RColorBrewer);
library(colorblindr); library(readxl);library(spatstat.utils);library(httr);library(cowplot)

source("dailyreport_dataprep_wk.R")

```

Here is the daily report for COVID-19 in Scotland for `r weekdays(max(scot_data$date))`, `r format(max(scot_data$date), '%d %B %Y')`. For the full dashboard, including comparisons with other countries, please visit [here](https://smazeri.shinyapps.io/Covid19_Scotland "COVID-19 Scotland").


**Deaths:**   
The death toll is currently `r comma(max(scot_deaths$deaths, na.rm = TRUE))`.   
The number of new deaths reported today was `r scot_deaths$new_deaths[nrow(scot_deaths)]`, which represents a `r formatC(100*(scot_deaths$deaths[nrow(scot_deaths)]-scot_deaths$deaths[nrow(scot_deaths)-1])/scot_deaths$deaths[nrow(scot_deaths)-1], format='f', digits=2 )`% increase on yesterday's total.    
The weekly ratio for new deaths measured over the last 7 days was `r formatC(wk_gr_newdeaths$ratio_m[nrow(wk_gr_newdeaths)], format='f', digits=2 )`. This is `r wk_gr_newdeaths$significance[nrow(wk_gr_newdeaths)]`. 

**Cases:**   
The total number of cases is currently `r comma(scot_data$confirmed_cases[nrow(scot_data)])`.    
The number of new cases today is `r scot_data$new_cases[nrow(scot_data)]`, which represents a `r formatC(100*(scot_data$confirmed_cases[nrow(scot_data)]-scot_data$confirmed_cases[nrow(scot_data)-1])/scot_data$confirmed_cases[nrow(scot_data)-1], format='f', digits=2 )`% increase on yesterday's total.    
The weekly ratio for new cases measured over the last 7 days was `r formatC(wk_gr_newcases$ratio_m[nrow(wk_gr_newcases)], format='f', digits=2 )`. This is `r wk_gr_newcases$significance[nrow(wk_gr_newcases)]`. 

**Hospital and ICU Occupancy**     
The number of patients in hospital with confirmed or suspected COVID-19 as of midnight is `r comma(filter(data_hosp, date == max(data_hosp$date))$Hospital_total)`, of whom `r filter(data_hosp, date == max(data_hosp$date))$ICU_total` are in ICU.    
The day-of-the-week ratio for hospital occupancy `r formatC(wk_gr_hosp_hosp_latest$ratio_m, format='f', digits=2 )`. This is `r wk_gr_hosp_hosp_latest$significance`.     
The day-of-the-week ratio for ICU occupancy is `r formatC(wk_gr_hosp_icu_latest$ratio_m, format='f', digits=2 )`. This is `r wk_gr_hosp_icu_latest$significance`.

```{r}
data_incidence <- scot_data_health_board %>%
  group_by(health_board) %>%
  arrange(date) %>%
  mutate(relative_increase_cases = 100*(confirmed_cases - replace_na(lag(confirmed_cases),0))/replace_na(lag(confirmed_cases),0)) %>%
  mutate(doubling_time_week = 7*log(2)/log(confirmed_cases/replace_na(lag(confirmed_cases,7),0))) %>%
  ungroup()


data_incidence_today <- subset(data_incidence, date == max(scot_data_health_board$date))

data_growth_rates_today <- subset(wk_gr_newcases_hb, date == max(wk_gr_newcases_hb$date)) %>%
                           arrange(desc(ratio_m)) %>% 
                           .[1,]
```


**Cases per capita (measured per 10,000 head of population):**   
The highest incidence over the last day was `r formatC(max(scot_ti_hb$today_incidence), format='f', digits=2 )`, which occurred in `r subset(scot_ti_hb, today_incidence == max(scot_ti_hb$today_incidence))$health_board`.   
The highest cumulative incidence is `r formatC(max(scot_ci_hb$cumulative_incidence), format='f', digits=2 )`, which is in `r subset(scot_ci_hb, cumulative_incidence == max(scot_ci_hb$cumulative_incidence))$health_board`.

**Health Boards\*:**   
The largest increase in cases in absolute terms by health board was `r max(data_incidence_today$new_cases)` cases, which occurred in `r subset(data_incidence_today, new_cases == max(data_incidence_today$new_cases))$health_board`.   
The largest relative increase in cases was `r formatC(max(data_incidence_today$relative_increase_cases),format='f', digits=2)`%, which occurred in `r subset(data_incidence_today, relative_increase_cases == max(data_incidence_today$relative_increase_cases))$health_board`.  
The highest weekly ratio for new cases was `r formatC(data_growth_rates_today$ratio_m, format='f', digits=2)`, which occurred in `r data_growth_rates_today$health_board`. `r data_growth_rates_today$sign_ci`.

**Tests:**   
The total number of tests reported is `r comma(scot_tests$Conducted[nrow(scot_tests)])`.   
The total number of positive cases to date is `r comma(scot_tests[nrow(scot_tests),"Total Positive"][[1]])`.   
The percentage of tests which were positive is currently `r formatC(round(100*scot_tests[nrow(scot_tests),"Total Positive"]/scot_tests$Conducted[nrow(scot_tests)],4)[[1]],format='f', digits=2 )`%.



\*To enhance graph readability and to account for cases being reallocated across health boards, the following health boards are grouped together: Grampian, Shetland and Orkney; Highland and Western Isles. 

The data in this report are provided by Health Protection Scotland and published daily by the Scottish Government [here](https://www.gov.scot/coronavirus-covid-19/ "gov.scot/coronavirus-covid-19/"). This report has been compiled in collaboration between members of the [Epigroup](https://www.wiki.ed.ac.uk/display/Epigroup/People "Epigroup website") and [EERA](https://www.ed.ac.uk/roslin/eeragroup "EERA website") groups at the University of Edinburgh. 
