---
title: "R Estimation - Scottish Outbreak"
author: "SM"
date: "03/05/2020"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tidyverse)
library(lubridate)
library(scales)
library(EpiEstim)
library(ggplot2)
library(incidence)
library(cowplot)
library(viridis)
library(knitr)

cases <- read_csv("scot_data.csv") %>%
  select(date, new_cases, confirmed_cases)

```

R estimation over the preceding 7 days using R package [EpiEstim](https://cran.r-project.org/web/packages/EpiEstim/vignettes/demo.html), based on daily cases reported.

Serial interval sources:

A) Mean 6.48 days and SD 3.83 as estimated in China (paper in Chinese). This is the estimate used by the Imperial College COVID-19 response team [here](https://mrc-ide.github.io/covid19-short-term-forecasts/index.html#methods) and by others [here](https://d212y8ha88k086.cloudfront.net/manuscripts/17379/dc06fb5c-fcf5-414d-bda8-8d4d61cea9b4_15843_-_kylie_ainslie.pdf?doi=10.12688/wellcomeopenres.15843.1&numberOfBrowsableCollections=5&numberOfBrowsableInstitutionalCollections=0&numberOfBrowsableGateways=13).  
B) Mean 3.96 days and SD 4.75 as estimated [here](https://wwwnc.cdc.gov/eid/article/26/6/20-0357_article).
C) Mean 4.7 days and SD 2.9 days as estimated [here](https://www.ncbi.nlm.nih.gov/pubmed/32145466).
D) Mean 7.4 and SD 5.2 days as estimated [here](https://arxiv.org/pdf/2003.05003.pdf).


```{r}
df <- data.frame(date = seq(min(cases$date), max(cases$date), "days"))
cases <- left_join(df, cases, by = "date") %>%
  mutate(new_cases = case_when(is.na(new_cases) ~ 0,
         TRUE ~ new_cases),
         confirmed_cases = zoo::na.locf(confirmed_cases))


## Estimate A Mean 6.48 and SD 3.83
res_parametric_si <- estimate_R(cases$new_cases, 
                                method="parametric_si",
                                config = make_config(list(
                                  mean_si = 6.48, 
                                  std_si = 3.83)))

df1<- res_parametric_si$R
df1$si <- 6.48
df1$date <- cases$date[8:nrow(cases)]

## Estimate B 3.96 days and SD of 4.75 estimated
res_parametric_si_3.96 <- estimate_R(cases$new_cases, 
                                method="parametric_si",
                                config = make_config(list(
                                  mean_si = 3.96, 
                                  std_si = 4.75)))

df2<- res_parametric_si_3.96$R
df2$si <- 3.96
df2$date <- cases$date[8:nrow(cases)]

## Estimate C Mean 4.7 days and SD 2.9 
res_parametric_si_4.7 <- estimate_R(cases$new_cases, 
                                     method="parametric_si",
                                     config = make_config(list(
                                       mean_si = 4.7, 
                                       std_si = 2.9)))

df3<- res_parametric_si_4.7$R
df3$si <- 4.7
df3$date <- cases$date[8:nrow(cases)]

## Estimate D Mean 7.4 and SD 5.2
res_parametric_si_7.4 <- estimate_R(cases$new_cases, 
                                     method="parametric_si",
                                     config = make_config(list(
                                       mean_si = 7.4, 
                                       std_si = 5.2)))

df4<- res_parametric_si_7.4$R
df4$si <- 7.4
df4$date <- cases$date[8:nrow(cases)]

df_all <- bind_rows(df1, df2, df3, df4)

nice_blue <- "#0072B2"
nice_red <- "#D55E00"
nice_green <- "#009E73"
nice_orange <- "#E69F00"


fig1 <- ggplot(filter(df_all, date >dmy("14/03/2020")), aes(x = date, y = `Mean(R)`, ymin = `Quantile.0.025(R)`, ymax = `Quantile.0.975(R)`, 
       fill = factor(si))) +
  geom_ribbon(alpha = 0.3) +
  geom_line(aes(colour = factor(si))) + 
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(breaks = c(0.5, 0.75, 1.0,2, 3,4,5)) +
  scale_x_date(date_breaks = "1 week", labels = date_format("%b %d")) +
  #scale_fill_manual(values = c(nice_blue, nice_red, nice_green, nice_orange)) + 
  #scale_colour_manual(values = c(nice_blue, nice_red, nice_green, nice_orange)) +
  scale_color_viridis(discrete = TRUE, option = "D")+
  scale_fill_viridis(discrete = TRUE) +
  labs(fill = "Serial \nInterval", colour = "Serial \nInterval", y = "R over the preceding 7 days", x = "Week Ending") +
  theme_bw() 

fig1_inset <- ggplot(filter(df_all, date > max(df_all$date)-8), aes(x = date, y = `Mean(R)`, ymin = `Quantile.0.025(R)`, ymax = `Quantile.0.975(R)`, 
       fill = factor(si))) +
  geom_ribbon(alpha = 0.2) +
  geom_line(aes(colour = factor(si))) + 
  geom_hline(yintercept = 1, linetype = "dashed") +
  scale_y_continuous(breaks = c(0.85, 0.9, 0.95, 1.0, 1.05, 1.1)) +
  scale_x_date(date_breaks = "2 days", labels = date_format("%b %d")) +
  #scale_fill_manual(values = c(nice_blue, nice_red, nice_green, nice_orange)) + 
  #scale_colour_manual(values = c(nice_blue, nice_red, nice_green, nice_orange)) +
  scale_color_viridis(discrete = TRUE, option = "D")+
  scale_fill_viridis(discrete = TRUE) +
  labs(fill = "Serial \nInterval", colour = "Serial \nInterval", y = "", x = "Week Ending") +
  labs( x = "",
        y = "",
        colour = "",
        fill = "") +
  theme_bw(base_size = 10) +
  theme(legend.position = "none") +
  theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))

ggdraw() +
  draw_plot(fig1) +
  draw_plot(fig1_inset, x = 0.4, y = .48, width = .40, height = .45)
```

\newpage 

```{r}
df_all %>%
  filter(date > max(df_all$date)-3) %>%
  select(date, si, `Mean(R)`, `Quantile.0.025(R)`, `Quantile.0.975(R)` ) %>%
  rename(Mean = `Mean(R)`,
         LCI = `Quantile.0.025(R)`,
         UCI = `Quantile.0.975(R)`) %>%
  arrange(desc(date)) %>%
  kable(caption = "Recent R estimates using published serial intervals", digits = 2)
```


